### 2022-09-06 manuscript/main.tex
initial commit

% set document class
\documentclass[a4paper,notitlepage,12pt]{article}

% load packages
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage[backend=biber,style=nature]{biblatex}

% specify bibliography source
\addbibresource{main.bib}
\renewcommand*{\bibfont}{\small}

% specify link style
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}

% specify some custom commands
\renewcommand{\baselinestretch}{1.2}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand*{\Affilfont}{\normalsize}

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1]{Samuel Zorowitz}
\author[1,2]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}

% specify the title
\title{Improving the reliability of cognitive task measures: A narrative review}

% turn date off
\date{}

% Define paragraph formatting
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

\maketitle

\section{Introduction}

% P1: Using valid and reliable measures of behavior is crucial for advancing biological psychiatry.

Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry. Here we define cognitive task as any experimental paradigm that measures some behavior in order to make inferences about one or more cognitive processes (e.g. Stroop task, delay discounting task, reversal learning task). Variability in performance on cognitive tasks provide experimenters with crucial insights into how individual differences in cognitive processing and ability relate to behavioral phenomena outside of the laboratory or clinic (e.g. reward sensitivity in a monetary-incentivized task and anhedonia in depression). Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{peng2019meta, spiegel2021relations} and cognitive decline \cite{salthouse2010selective}. Cognitive tasks then are invaluable tools for building and refining our understanding of psychiatric symptoms and syndromes. For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.

% P2: Three key properties of behavioral measures: discriminability, reliability, and validity

The psychometric quality of a cognitive task measure (i.e. any descriptive or model-based indicator of behavioral performance such as proportion correct, average response time, and learning rate) for characterizing individual differences are summarized by three key properties: discriminating power, reliability, and validity \cite{kline2015handbook}. The discriminatory power of a task measure describes its ability to produce variability in participants' performance. This is a necessary property of tasks to be used for studying individual differences; where there is no variation in performance, there are no individual differences to study. The reliability of a task characterizes the degree to which it consistently rank-order participants across measurements. Put another way, a task measure is reliable if, assuming participants have not changed, it produces the same score for each participant over time. Finally, the validity of a task measure concerns whether it actually measures what it purports to measure. The focus of the current review is task reliability, which is a prerequisite for validity.

% P3: definition and Importance of reliability

What is the formal definition of reliability? In classical test theory, the variance in observed scores on a task measure (X) is the sum of the true score variance (T), reflecting real individual differences in the latent construct of interest, and measurement error (E; i.e. X = T + E). The reliability of a measure is defined as the proportion of variance attributable to the true score variance relative to total variance (i.e. R = T/[T + E]). In other words, reliability quantifies the magnitude of individual differences relative to the noisiness of a task measure; the larger the reliability of a task measure, the more it reflects true individual differences and the less it reflects noise.. It should now be clearer why reliability is a prerequisite for validity; when a task measure is unreliable, it reflects measurement error and not the construct of interest. If this were not reason enough to care about reliability, the observed correlation between two measures (e.g. task measure and self-reported symptom score) is a function of their individual reliabilities \cite{Spearman1904-mo}:

\begin{equation}
    \rho_{xy} = \rho_{tt} \sqrt{\rho_{xx'} \cdot \rho_{yy'}}
\end{equation}

where $\rho_{xx'}$ and $\rho_{yy'}$ are the reliabilities of two measures, $x$ and $y$; $\rho_{tt}$ is their true latent correlation; and  $\rho_{xy}$ is their observed correlation. As can be observed in the equation above, the reliability of a measure places an upper bound on the maximum observable correlation between itself and a second measure. As reliability decreases, so too does the maximum observable correlation between itself and any other measure (Figure 1). As an important corollary, as reliability decreases, the number of participants required to reliably detect a correlation between two measures increases \cite{Parsons2019-jw}. Thus, poor reliability hampers our ability to investigate associations between cognitive processes, as measured by task performance, and other variables of interest.

% P4: Reliability is inherently context-dependent: it is a function of the population, task, and testing context

To further complicate matters, the reliability of a task measure is not universal. The reliability of a task reflects interactions between its design, the participants completing it, and the context in which it is administered. Indeed task reliability can vary as a function of experimental design (stimulus sets, number of trials, time limits \cite{paap2016role, cooper2017role}); sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in-clinic, online); and response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); and scoring or estimation method \cite{Rouder2019-am, haines2020learning}. For example, a cognitive task originally designed for use with an adult population may prove too difficult for children; their task performance may drop to chance-level, thereby minimizing between-participant variance and, as a consequence, task reliability. As a second example, participants completing an experiment online from their homes may experience more distraction than if they participated in the lab. This may in turn lead to an increase in measurement error and a concomitant decrease in reliability. Experimenters therefore cannot assume the reliability of a task measure is constant. Ideally then, experimenters should investigate and report the reliability of task measures as part of any analysis or, at the very least, after having made changes to a task or scoring procedure or when administering the task to new sample populations or in new testing contexts.

% P5: Task reliability lower on average

Though verifying the reliability of cognitive task measures is paramount to successful individual differences research, the reliability of task measures are historically seldom reported \cite{Green2016-xw, Parsons2019-jw}. When reported, it has frequently been observed that task measures exhibit lower reliability than what is conventionally considered acceptable for individual differences research (e.g. $\rho$ = 0.7 - 0.8) and lower than what is regularly achieved by self-report measures. Indeed many studies have now found that task measures to exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}. One possible explanation for this finding is the so-called ``reliability paradox'' of cognitive tasks \cite{hedge2018reliability}, which states that the often lackluster reliability of tasks is a result of a mismatch in goals between experimental and individual differences psychological research. In experimental psychology, the goal is often to demonstrate the existence of a behavioral effect. One sure-fire means of increasing the power to detect an effect is to minimize between-participants variance of a task measure. This of course is the exact opposite of what is desirable for individual differences research, where between-participants variance is essential to achieving reliable task measures. Thus, the tendency in biological psychiatry to adopt the most prominent tasks in experimental psychology --- the ones that most reliably demonstrate a behavioral effect --- may actually hamstring our efforts to study individual differences.

% P6: Purpose of paper: narrative review of task reliability

Regardless we do not believe that task measures are inherently less reliable and that pessimism about task-based individual differences research is warranted. It is possible to (re)design tasks to achieve good reliability, even to the high levels dictated by conventional standards \cite{waltmann2022sufficient, sullivan2022enhancing, kucina2022solution, snijder2022psychometric}. The purpose of the current article is to provide a narrative review of approaches to improve task measure reliability. The article is divided into two main parts. First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks. Then, we introduce a taxonomy of approaches for improving task reliability using concrete examples from the published literature. We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual differences research.

\section{Calculating reliability}

We begin with a brief discussion of calculating reliability for cognitive task measures. There are two measures of reliability of relevant for task measures: internal consistency and test-retest reliability. Internal consistency is the reliability of a measure in a single administration. In turn, test-retest reliability (also known as temporal stability) is reliability of a task measure across two or more administrations. The two types of reliability provide complimentary information: internal consistency quantifies the consistency of individual difference on a task measure in a given testing session, whereas test-retest reliability quantifies the stability of individual differences across time. Even for experimenters primarily interested in cross-sectional experiments, test-retest reliability is a useful index. This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant state-dependent factors (e.g. current mood, fatigue). Given sufficient time between testing sessions, test-retest reliability should be less biased state factors. A second point to consider with respect to these two indices is that, whereas high internal consistency is virtually always desirable for individual difference studies, high test-retest reliability may not be. Test-retest reliability should be interpreted with respect to the construct of interest. For example, low 1-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g. mood effects on cognition), but may be problematic if the effect is hypothesized to be stable across time (e.g. personality effects on cognition).

The test-retest reliability of a task measure can be calculated in numerous ways \cite{Parsons2019-jw}. Perhaps the simplest approach is to compute the Pearson correlation between participants' scores from two sessions. An alternative approach is to calculate the intraclass correlation coefficient (ICC), which decomposes a task measure into true score variance and error variance. There are many formulas for calculating ICC \cite{Koo2016-fz}, with the critical distinction being reliability is based on the \textit{consistency} or \textit{absolutely agreement} of a task measure across two administration. Consistency-based ICCs are affected only by the relative ordering of participants across time; that is, they are insensitive to systematic changes to the actual values of a task measure across time (e.g. practice effects). In contrast, absolute agreement-based ICCs measure the degree to which scores are identical across time. They are sensitive to changes in the overall mean of a task measure across sessions. The type of ICC to use depends on the experimenter's goals. If an experimenter cares only about the rank-ordering of participants across time, a consistency-based ICC (or Pearson correlation) is preferable. Conversely, if systematic differences in a task measure across task administrations is a concern, then an absolute agreement-based ICC should be preferred.

Calculating the internal consistency is more complicated for task measures. The most common measure of internal consistency, Cronbach's $\alpha$, requires certain assumptions that are unrealistic for task data (e.g. equivalence of trials, uncorrelated measurement error). As such, internal consistency for task measures is instead usually calculated via split-$K$ reliability instead, where $K$ indicates the number of splits. Split-half reliability ($K=2$) is most frequently chosen. A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be biased when partitions are not equally good measures of a behavior \cite{Green2016-xw, pronk2022can}. For example, first-second splitting (i.e. partitioning the data by the first and second halves of an experiment) may be confounded by practice, fatigue, or other linear time effects. Under such conditions, first-second partitioning will spuriously decrease the similarity of data across partitions, thereby increasing estimates of measurement noise and underestimating reliability. In contrast, odd-even splitting (i.e. partitioning the data by odd and even trials) may be confounded when behavior across trials is non-independent (i.e. correlated errors). Under such conditions, odd-even splitting will artificially inflate the similarity of data across partitions, thereby decreasing estimates of measurement noise and overestimating reliability. Therefore, where possible, a permutation-based approach to calculating split-half reliability has been recommended \cite{pronk2022can, Parsons2019-jw}. Briefly, the idea is to compute the average reliability across many thousands of random partitions of the data into halves. For task measures derived from cognitive models, it may be unreasonably computationally intensive to employ such an approach. In which case, it may instead be preferable to design tasks with at least two independent blocks where the model is fit to each block independently and reliability is calculated using the model parameters estimated from each.

A final point is that, when calculating reliability for task measures, traditional sum or mean score estimates of performance (e.g. proportion correct responses, mean response time) may substantially underestimate reliability \cite{Rouder2019-am, haines2020learning}. These summary scores of task performance neglect trial-by-trial variability in performance, which can result in noisier estimates of task performance and consequently diminished estimates of reliability. Instead, it is often preferable to use trial-level hierarchical (mixed effects) models that can simultaneously estimate between- and within-participant (i.e. trial-level) variability in behavior. By directly modeling trial-by-trial variability, estimates of participants' performance are effectively denoised. Furthermore, the use of hierarchical models allows for partial pooling of trial data across participants, which further increases the precision of estimates of participants' performance. Together these properties can substantially boost estimates of reliability \cite{Rouder2019-am, haines2020learning}. In practice, multiple studies have demonstrated how the use of hierarchical trial-level models directly improves estimates of task measure reliability \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}. For a detailed discussion of hierarchical models in the context of task reliability, see Haines (this issue).

\section{Improving task reliability}

As defined above, the reliability of a task measure is proportion of variance attributable to between-participant differences relative to measurement error. Thus, there are two major strategies for improving the reliability of a measure: by increasing between-participant variability or by decreasing measurement error. In what follows, we discuss approaches for accomplishing both strategies in turn. Where appropriate, we highlight studies from the published literature that are exemplary for imporiving the reliability of a task measure by implementing a particular strategy.

\subsection{Increasing between-participant variance}

\subsubsection{Ceiling \& floor effects}

By definition, the reliability of a task measure is zero when there is no variability across participants. Thus, range restriction of task measures via ceiling or floor effects is a serious obstacle to reliability. Siegelman and colleagues \cite{siegelman2017measuring} noted the consequences of floor effects on reliability in the context of statistical learning tasks. In statistical learning tasks, participants must learn to identify subtle patterns in the transition probabilities underlying a continuous sequence of stimuli. In reanalyzing archival datasets, Siegelman and colleagues found a majority of participants were at chance-level performance in discriminating between legitimate and foil sequence patterns; consequently, the reliability of conventional proportion correct measures suffered. In response, the authors designed a new version of a statistical learning task involving stimulus sequences that ranged more widely in their difficulty to learn. Only a minority of participants were at chance-level performance on this new task and, as such, the reliability of proportion correct scores improved. Similarly, in developing an abbreviated working memory task, Oswald and colleagues \cite{oswald2015development} found that they could remove the easiest trials --- those with ceiling level performance --- with virtually no change to task reliability. This is because working memory trials with performance at ceiling are incapable of differentiating ability across participants and therefore cannot contribute significantly to the reliability of the task.

Experimenters who are administering a task to a new population should be wary of range restriction effects. Cognitive tasks calibrated for one group of participants may be too easy or difficult for a different group. For example, Arnon and colleagues \cite{arnon2020current} found that statistical learning tasks developed for adults were too difficult for young children and therefore yielded unreliable discrimination scores in that population. Similarly, Kyllonen and colleagues \cite{kyllonen2019general} were motivated to develop a battery of fluid reasoning measures for highly educated adults due to observed ceiling effects in performance when using preexisting fluid reasoning tasks in this population. In sum, experimenters should be cautious when administering preexisting tasks to new populations; the tasks calibrated for one group may not be adequately sensitive for others.

\subsubsection{Repeatability \& practice effects}

A related issue for task reliability is practice effects, or when participants' performance on a task improves with repeated administrations. Practice effects are relatively common for cognitive tasks \cite{hausknecht2007retesting, scharfen2018retest}. They might occur due to the attenuation of task-irrelevant nuisance factors (e.g. performance anxiety) and/or the learning of task-specific knowledge or strategies. Practice effects are not inherently an issue for reliability --- especially if an experimenter is only interested in the consistency, but not the absolute agreement, of participants' performance over time --- but they can become a pernicious issue if they are severe enough to induce ceiling effects. For example, Paredes and colleagues \cite{paredes2021psychometric} observed large practice effects on the Pavlovian go/no-go task for short retest intervals (3 days, 14 days), which resulted in poor estimates of test-retest reliability. In developmental and lifespan studies, practice effects are potentially complicated by their interaction with age \cite{anokhin2022age, salthouse2010influence}; that is, practice-induced ceiling effects may present in some age groups but not others.

One strategy for minimizing practice effects is simply to increase the time interval between task administrations. The more time that elapses between sessions, the greater the probability that participants have forgotten task-specific knowledge or strategies \cite{hausknecht2007retesting, scharfen2018retest}. Of course, this solution may not always be possible or desirable, especially if an experimenter is only able to or specifically interested in studying a behavior over a short time period.

A second strategy is to design tasks so as to prevent or discourage the formation of task-specific strategies. For example, McLean and colleagues \cite{mclean2018towards} investigated the repeatability of the beads task. In the beads task, participants are presented with two jars containing beads of two colors in equal but opposite ratios. In one trial of the task, a predetermined sequence of beads is drawn from one jar. After participants must decide which jar beads were being drawn from or request to see more beads. In a typical version of the task, the same sequence of bead draws is used across all trials. McLean and colleagues found that participants were aware that the sequence repeated across trials and, as a consequence, became more erratic in their decision to witness more bead draws with additional trials. In response, the authors developed a new version of the beads task that included distractor sequences of bead draws. This new design prevent participants from becoming aware of the target sequence, which in turn motivated participants to respond more consistently thereby improving the reliability of participants' information seeking scores.

\subsubsection{Enhancing experimental effects}

The preceding sections described potential threats to between-participants variability, but did not address how to actually increase it. A primary strategy for increasing between-participants variability (and therefore reliability) is to enhance the experimental effect. Amplifying the magnitude of an experimental effect (e.g. making a task more challenging, increasing the potency of a stress or affective induction) typically increases the range of participants' responses to it. Two recent studies highlight this approach in the context of cognitive control tasks.

Kucina and colleagues \cite{kucina2022solution} investigated the reliability of cognitive conflict effects in new versions of several standard cognitive control tasks (e.g. Stroop, Flanker, Simon) that amplified cognitive interference via two task design features. First, they combined multiple sources of cognitive interference in the same task (e.g. combining the Stroop and Simon effects to create a ``Stroopon'' task). Second, for a subset of trials, they required participants to make multiple responses based on both relevant and irrelevant stimuli attributes. Compared to the standard version of these tasks, these experimental manipulations had the effect of increasing task demands, which resulted in greater between-participants variance and reliability as a consequence. Similarly, Snijder and colleagues \cite{snijder2022psychometric} studied the reliability of cognitive conflict effects in multiple redesigned cognitive control tasks (e.g. Stroop, AX-CPT, Cued Task Switching, Sternberg) that required of participants either more proactive control (i.e. maintained vigilance) or reactive control. The authors found that the proactive control versions of the task were more demanding, which similarly led to larger cognitive interference effects and ultimately task reliability.

For decision-making tasks, or other tasks requiring participants to choose between two or more discrete options, a related strategy is to calibrate the difficulty of trials to the average ability of the population of interest. For example, consider a task trial with two response options. Given the Bernoulli distribution, the variance in responses on this trial is maximized when the probability of choosing either response is equal (i.e. 50\%/50\%). Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items are around the average ability of the sample \cite{gulliksen1945relation, ferrando2007external} (or somewhere above this point when guessing the correct response is an issue \cite{lord1952relation, feldt1993relationship}). Of course, this design principle is only helpful to the degree that an experimenter  precisely knows the average ability level of their participants. If this is unknown or poorly characterized, then it is instead desirable to design a task with trials spanning a uniform range of difficulty.

\subsubsection{Sample population}

A final strategy for increasing between-participant variance is to simply recruit more diverse samples. While convenient, undergraduate students from a single university are likely to be relatively homogeneous in their cognitive profiles. It may be worthwhile instead to recruit participants from the community or from an online labor market (e.g. Amazon Mechanical Turk, Prolific Academic, CloudResearch Panels). With regard to the latter, it is worth mentioning that, because online participants are typically completing experiments from their homes, they may be more likely to be distracted or multi-tasking during an experiment \cite{newman2021data}. Thus, when recruiting online samples, experimenters should take special care to ensure that an increase in between-participants variance is not offset by a concomitant increase in measurement noise.

\subsection{Decreasing measurement noise}

\subsubsection{Increasing trial numbers}

Perhaps the most straightforward approach to decreasing measurement error, and thereby increasing reliability, is to increase the number of task trials. In generalizability theory, the relationship between reliability and the number of trials defined as:

\begin{equation}
    \rho = \frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E / n}
\end{equation}

where $\sigma^2_T$ is the true between-participants variance, $\sigma^2_E$ is measurement error (i.e. trial-level variance), and $n$ is the number of trials. In practice, this relationship often holds (e.g. \cite{paap2016role, cooper2017role}) but there are sometimes exceptions \cite{price2015empirical, klingelhoefer2022robust}. Increasing the number of task trials only benefits reliability if measurement error is random. If increasing task length results in participant fatigue or boredom, then measurement noise may systematically increase and, as a consequence, reliability will suffer. Furthermore, even if increasing the trial number does not induce fatigue, it may still be impractical for other reasons. Because increasing the trial number yields diminishing marginal improvements for reliability, achieving a desired level of reliability may require prohibitively long experiments, especially when the task is completed as part of a larger battery.

\subsubsection{Improving experiment designs}

Measurement error can be reduced through improving the design of experiments, which can be accomplished in many ways. The reliability of a task measure can be improved by including in an experiment only the most discriminating stimuli. For example, in the context of an emotion recognition task, stimuli of good discriminability with would be those where participants with good emotion recognition ability consistently correctly identify the displayed emotion and were participants with poor ability consistently incorrectly identify the displayed emotion \cite{keutmann2015generating}. In contrast, stimuli with poor discriminability --- those for which performance between high and low ability participants is indistinguishable --- will lead to more measurement noise and decreased reliability. In experiments where stimuli are intended to be unique and distinguishable, it is important to ensure that participants are not confusing the identities of the stimuli. Improving both the linguistic and visual distinctness of stimuli may prevent this sort of confusion and therefore aid reliability \cite{yoo2022importance}.

Other design features of an experiment may affect reliability. Consider for example the dot-probe task, in which participants must disengage attention from a distracting image on one part of the screen in order to identify and respond to the orientation of a pair of dots elsewhere on the screen. Previous research has found that dot-bottom trials, where a participant disengage from a distracting stimulus located at the top of the screen and then saccade to the bottom of the screen, are more reliable than dot-top trials \cite{price2015empirical, aday2019extended}. An explanation for this finding is that, if participants tend to have a gaze bias towards the top of the screen, a saccade away from the top of the screen might require a stronger level of disengagement. As such, dot-bottom trials are sometime found to be a more robust signal of attentional bias. In summary, the actual mechanics of a task trial may have profound influence on task reliability

It is worth stressing that clear, concise instructions are essential for task reliability. When participants are unsure of what they are intended to do in an experiment, their behavior is likely to be more erratic and random, thereby diminishing reliability. (Clear instructions may also work to ensure the validity of an experiment by discouraging participants from using cognitive strategies not of interest to the experimenter.) In their 10 simple rules paper for designing cognitive experiments, Barbosa and colleagues \cite{barbosa2022practical} provide practical suggestions for writing task instructions. When adapting a task for a new population, experimenters should ensure that the instructions are still appropriate. Task instructions that are comprehensible to healthy adult participants may not be suitable for other populations like children \cite{hughes2002measuring}.

Another strategy to reduce measurement error is to incorporate practice trials into an experiment. Practice trials can help to minimize the effects of nuisance factors such as performance anxiety or unfamiliarity with the response modality. Moreover, practice trials give participants participants an additional opportunity to make sure they understand the task instructions. Thus, practice trials can help participants reach a more consistent ``steady state'' of behavior, thereby reducing the noisiness of their responses and increasing reliability \cite{alexander2003effects}. If experimenters are unable to include a standalone practice block in their experimenters due to time constraints, they can implicitly incorporate practice trials by designating and discarding (or modeling separately) the first handful of trials during an experiment \cite{mclean2018towards}.

Yet another strategy to diminish measurement error is to ``gamify'' an experiment. The gamification of cognitive tasks, --- incorporating (video) game design elements into cognitive tasks --- can increase participant engagement and motivation during an experiment \cite{sailer2017gamification}. Keeping participants motivated and engaged may counter the would-be effects of boredom and fatigue on task reliability. For example, Kucina and colleagues \cite{kucina2022solution} cite task gamification as an important factor that contributed to the improved the reliability of their cognitive control tasks. Similarly, Verdejo and colleagues \cite{verdejo2021unified} partially attribute the improved reliability of their impulsivity task battery to gamified task design.

\subsubsection{Reducing parameter estimation noise}

When the parameters from cognitive models are used as summaries of participants' task performance, another means to improve measure reliability is to decrease estimation noise. The estimation noise of a parameter given an experiment and model can be quantified through simulation studies \cite{wilson2019ten, palminteri2017importance}. In a simulation study, an experimenter generates artificial data for the experiment using representative model parameters. The experimenter attempts to recover the model parameters by fitting the model to the simulated data. Estimation noise is the inverse of the (relative or absolute) agreement between the true and recovered parameters. Alterations to experimental design can improve parameter recovery and estimation noise, and multiple frameworks have been proposed for testing and improving experimental designs to aid parameter recovery \cite{broomell2014parameter, melinscak2020computational}. Parameter recovery can also be affected by model estimation method \cite{Lerche2017-yg, waltmann2022sufficient}. As mentioned above, the partial pooling properties of hierarchical Bayesian models can be especially beneficial for improving parameter recovery and decreasing estimation noise \cite{katahira2016hierarchical}.

A related approach is to use adaptive experimental designs \cite{myung2013tutorial}. In an adaptive experiment, the design of an experiment is generated on-line in order to present each participant with stimuli or trial types that are matched to their particular response patterns or ability levels. Though undoubtedly a more complex experimental design, adaptive experiments have the advantage of selecting the most informative trials for resolving the ability or preference level of a participant (e.g. as measured by a cognitive model). Furthermore, adaptive designs  may require fewer trials to achieve precision measurement, which can save experimenters (and participants) time and money. Adaptive designs have been successfully translated in cognitive research for example in working memory \cite{gonthier2018measuring} and delay discounting \cite{ahn2020rapid}. For a detailed discussion of adaptive design experiments, see Ahn (this issue).

Parameter estimation may be further improved by leveraging additional information. For example, latent variables may be more accurately measured through the inclusion of demographic variables or other covariates during estimation \cite{anders2018improved, Curran2016-bt}. The basic idea is that, if there are any associations between model parameters and covariates, they can aid in resolving parameter estimates. An extension of this idea is to utilize joint modeling of dependent variables; that is, to design models where multiple observed trial-level variables are predicted simultaneously. For example, the joint modeling of choice and response time has been found to improve the precision and reliability of estimated parameters in cognitive ability testing \cite{bertling2018using} and reinforcement learning \cite{ballard2019joint, shahar2019improving}. It is also possible to incorporate physiological and/or neural correlates of behavior, such as skin conductance response, fMRI BOLD signal, and EEG.

\subsection{Difference scores}

In the context of reliability, difference scores deserve special treatment. Difference scores describe a measure where a measure of a participant's performance from one condition is subtracted from their performance in another. An example of a difference score is the classic Stroop interference effect; that is, the difference in average reaction time between congruent and incongruent trials. The value of difference scores is straightforward: difference scores allow experimenters to isolate particular cognitive processes (e.g. processing cognitive conflict) and to control for irrelevant sources of variance (e.g. perceptual processing, motor ability). Given their utility, difference scores are ubiquitous in experimental psychology.

The challenge for difference scores is that their reliability is a function of the reliability of their components \textit{and} the correlation between the components:

\begin{equation}
    \rho_{dd'} = \frac{\sigma^2_x \rho_{xx'} + \sigma^2_y \rho_{yy'} - 2 \rho_{xy} \sigma_x \sigma_y}{\sigma^2_x + \sigma^2_y - 2 \rho_{xy} \sigma_x \sigma_y}
\end{equation}

where $\sigma^2_x$ and $\sigma^2_y$ are the variances of task measures \textit{x} and \textit{y}; $\rho_{xx'}$ + $\rho_{yy'}$ are the reliabilities of task measures \textit{x} and \textit{y}; and $\rho_{xy}$ is the correlation between task measures \textit{x} and \textit{y} \cite{trafimow2015defense, chiou1996reliability}. When the variances of the two measures are equal, this reduces to:

\begin{equation}
    \rho_{dd'} = \frac{\rho_{xx'} + \rho_{yy'} - 2 \rho_{xy}}{2 - 2 \rho_{xy}}
\end{equation}

From the above equation, it is readily apparent that the reliability of a difference score measure is diminished to the extent that its components are correlated. In experimental psychology, two measures derived from the same task will often be correlated due to shared domain-general cognitive processes. Thus, it will often (if not always) be the case that difference scores derived from task measures will be less reliable than their components. This presents a challenge for designing reliable difference score measures.

\subsubsection{Identify alternative measures}

An alternative approach to improving the reliability of difference scores is simply to avoid using them in the first place. That is, experimenters may opt to design tasks or analyses in such a way to avoid the necessity of difference scores in the first place. This recommendation has a long history in experimental psychology. Indeed, because difference scores will virtually always be less reliable, many authors have advocated to simply abandon them \cite{lord1956measurement, cronbach1970we, edwards2001ten}.

What then are the alternatives to difference scores? Draheim and colleagues \cite{draheim2019reaction} provide an in-depth review of alternatives to difference scores in the context of response time measures, though much of their discussion is applicable to difference scores in general. One possibility is to simply use the component measures (e.g. performance on incongruent trials in a Stroop task alone). Of course, because component scores will be contaminated other sources of variance, such as baseline performance, interpreting component scores should be done with caution. Another approach is to identify alternative measures of task performance. For example, intra-individual response time variability has been identified as a correlate in executive control, with aberrations in psychopathology broadly \cite{kofler2013reaction}, that can be measured reliably \cite{saville2011stability}.

\subsubsection{Enhance \& purify task measures}

When difference scores are unavoidable, there still steps experimenters can take to improve their reliability. As the equations above suggest, the reliability of difference scores can be improved in three ways: (1) improve the reliability of the component measures, (2) increase the differential variability of the task measures, and (3) minimize the correlation between the task measures. The first strategy has been the focus of the discussion so far. The second approach deserves further comment. As noted elsewhere \cite{trafimow2015defense, chiou1996reliability}, the reliability of a difference score measure increases as ratio of measure variances increases. Figure XX shows that, even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the deviation ratio is high. This speaks to another advantage of increasing the between-participants variability of a measure (insofar that an experimenter increases one, but not the other, measure). This result might help explain the improved reliability of the difference score measures studied by Kucina and colleagues \cite{kucina2022solution}.

The third approach is to purify task measures; that is, to decorrelate the components of a difference score by reducing or eliminating their shared variance. Rey-Mermet and colleagues \cite{rey2019executive} provide an interesting example in the context of executive control. In typical executive control tasks, response times on congruent and incongruent trials are highly correlated, reflecting shared variance from conflict-irrelevant processes including baseline processing speed (e.g. perceptual processing, motor speed) and performance strategies (e.g. individual differences in speed-accuracy preferences; \cite{draheim2019reaction}). Noting this, Rey-Mermet and colleagues designed a number of ``response-deadline'' executive control tasks, where participants had a limited amount of time to respond during a trial. Specifically, the duration of the response deadline was calibrated for each participant individually in order to achieve a fixed accuracy level in blocks of neutral trials, which was then used as the deadline for both congruent and incongruent trials. They reasoned the calibration procedure had several important advantages. First, it controls for individual differences in processing speed that contribute to performance in both congruent and incongruent trials. Second, it also controls for individual differences in strategy, because, regardless of whether a participant was biased towards speed or accuracy, inefficient executive control would result in lower accuracy. They found that, with the calibration procedure, the reliability of an accuracy difference score (incongruent - congruent) was as good or better than what had previously been reported for executive control tasks. Thus, controlling for shared variance across measures --- that is, purifying measures --- can help to improve task reliability.

\section{Discussion}

\printbibliography

\end{document}

### 2022-10-26 manuscript/main.tex
checkpointing progress

@@
 Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry.
-Here we define cognitive task as any experimental paradigm that measures some behavior in order to make inferences about one or more cognitive processes (e.g.
-Stroop task, delay discounting task, reversal learning task).
-Variability in performance on cognitive tasks provide experimenters with crucial insights into how individual differences in cognitive processing and ability relate to behavioral phenomena outside of the laboratory or clinic (e.g.
+Variability in performance on cognitive tasks provides experimenters with crucial insights into the relationship between individual differences in cognitive processing and behavioral phenomena outside of the laboratory or clinic (e.g.
 reward sensitivity in a monetary-incentivized task and anhedonia in depression).
-Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{peng2019meta, spiegel2021relations} and cognitive decline \cite{salthouse2010selective}.
-Cognitive tasks then are invaluable tools for building and refining our understanding of psychiatric symptoms and syndromes.
+Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{spiegel2021relations} and cognitive decline \cite{hartshorne2015does}.
+Cognitive tasks, then, are invaluable tools for building and refining our understanding of psychiatric symptoms and syndromes.
 For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.
 
-% P2: Three key properties of behavioral measures: discriminability, reliability, and validity
+% P2: Definitions + Three key properties of behavioral measures: discriminability, reliability, and validity
 
-The psychometric quality of a cognitive task measure (i.e.
-any descriptive or model-based indicator of behavioral performance such as proportion correct, average response time, and learning rate) for characterizing individual differences are summarized by three key properties: discriminating power, reliability, and validity \cite{kline2015handbook}.
-The discriminatory power of a task measure describes its ability to produce variability in participants' performance.
+We define a cognitive task as any experimental paradigm that measures behavior in order to make inferences about one or more cognitive processes (e.g.
+Stroop task, delay discounting task, reversal learning task).
+Cognitive-task measures for characterizing individual differences are descriptive or model-based indicators of behavioral performance such as proportion correct, average response time, and learning rate.
+The psychometric quality of such a measure can be summarized by three key properties: \emph{discriminating power}, \emph{validity} and \emph{reliability} \cite{kline2015handbook}.
+The discriminatory power of a task measure describes its ability to measure variability in participants' performance.
 This is a necessary property of tasks to be used for studying individual differences; where there is no variation in performance, there are no individual differences to study.
-The reliability of a task characterizes the degree to which it consistently rank-order participants across measurements.
+The validity of a task measure concerns whether it actually measures what it purports to measure.
+Finally, the reliability of a task measure characterizes the degree to which it consistently rank-orders participants across measurement instances.
 Put another way, a task measure is reliable if, assuming participants have not changed, it produces the same score for each participant over time.
-Finally, the validity of a task measure concerns whether it actually measures what it purports to measure.
-The focus of the current review is task reliability, which is a prerequisite for validity.
+The focus of this review is task reliability, which is a prerequisite for validity.
 
 % P3: definition and Importance of reliability
+\subsubsection*{The formal definition of reliability}
 
-What is the formal definition of reliability?
-In classical test theory, the variance in observed scores on a task measure (X) is the sum of the true score variance (T), reflecting real individual differences in the latent construct of interest, and measurement error (E; i.e.
-X = T + E).
-The reliability of a measure is defined as the proportion of variance attributable to the true score variance relative to total variance (i.e.
-R = T/[T + E]).
-In other words, reliability quantifies the magnitude of individual differences relative to the noisiness of a task measure; the larger the reliability of a task measure, the more it reflects true individual differences and the less it reflects noise..
-It should now be clearer why reliability is a prerequisite for validity; when a task measure is unreliable, it reflects measurement error and not the construct of interest.
+In classical test theory, the variance in observed scores on a task measure $X$ is the sum of the true score variance $T$, reflecting real individual differences in the latent construct of interest, and measurement error $E$, i.e., $X = T + E$.
+The reliability of a measure is defined as the proportion of variance attributable to the true score variance relative to total variance: $\rho = \frac{T}{T + E}$.
+Thus, reliability quantifies the magnitude of individual differences relative to the noisiness of a task measure; the larger the reliability of a task measure, the more it reflects true individual differences rather than reflecting noise.
+Reliability is therefore a prerequisite for validity: an unreliable task measure reflects measurement error and not the construct of interest.
 If this were not reason enough to care about reliability, the observed correlation between two measures (e.g.
-task measure and self-reported symptom score) is a function of their individual reliabilities \cite{Spearman1904-mo}:
-
+task performance and self-reported symptom score) is bounded by their individual reliabilities \cite{Spearman1904-mo}:
 \begin{equation}
@@
 \end{equation}
+where $\rho_{xx'}$ and $\rho_{yy'}$ are the reliabilities of two measures, $x$ and $y$; $\rho_{tt}$ is their true latent correlation; and  $\rho_{xy}$ is their observed correlation.
+As all reliabilities are $<1$, the reliability of a measure places an upper bound on the maximum observable correlation between itself and a second measure (Figure \ref{fig:fig01}).
+As an important corollary, as measure reliability decreases, the number of participants required to reliably detect a correlation between two measures increases \cite{Parsons2019-jw}.
+Thus, poor reliability hampers our ability to investigate associations between cognitive processes, as measured by task performance, and other variables of interest.
 
-where $\rho_{xx'}$ and $\rho_{yy'}$ are the reliabilities of two measures, $x$ and $y$; $\rho_{tt}$ is their true latent correlation; and  $\rho_{xy}$ is their observed correlation.
-As can be observed in the equation above, the reliability of a measure places an upper bound on the maximum observable correlation between itself and a second measure.
-As reliability decreases, so too does the maximum observable correlation between itself and any other measure (Figure 1).
-As an important corollary, as reliability decreases, the number of participants required to reliably detect a correlation between two measures increases \cite{Parsons2019-jw}.
-Thus, poor reliability hampers our ability to investigate associations between cognitive processes, as measured by task performance, and other variables of interest.
+\begin{figure}[tp]
+    \centering
+    \includegraphics[width=1.0\textwidth]{figures/fig01.png}
+    \caption{\small The relationship between measure reliability, observed correlations, and statistical power.
+(A) The maximum expected observed correlation between two measures as a function of their true (latent) correlation and reliability.
+As the reliability of two measures decreases, so too does their observed correlation.
+(B) Required sample size for 80\% power to detect true correlations between two measures given their reliability.
+As the reliability of two measures decreases, the number of participants required to detect an association increases.
+Even with a large sample size of $N=500$, a measure with reliability $\rho_{tt} = 0.5$ will only allow detection of correlations above $0.3$, which are uncharacteristically high in the field.}
+    \label{fig:fig01}
+\end{figure}
 
@@
 
-To further complicate matters, the reliability of a task measure is not universal.
-The reliability of a task reflects interactions between its design, the participants completing it, and the context in which it is administered.
-Indeed task reliability can vary as a function of experimental design (stimulus sets, number of trials, time limits \cite{paap2016role, cooper2017role}); sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in-clinic, online); and response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); and scoring or estimation method \cite{Rouder2019-am, haines2020learning}.
-For example, a cognitive task originally designed for use with an adult population may prove too difficult for children; their task performance may drop to chance-level, thereby minimizing between-participant variance and, as a consequence, task reliability.
+To further complicate matters, the reliability of a task measure is not absolute -- it reflects interactions between task design, the participants, and the context in which the task is administered.
+Indeed, task reliability can vary as a function of parameters of the experiment such as the specific stimulus set, number of trials, time limits \cite{paap2016role, cooper2017role}); sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in clinic, online); response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); and scoring or estimation method \cite{Rouder2019-am, haines2020learning}.
+For example, a cognitive task originally designed for use with an adult population may prove too difficult for children.
+Their task performance may drop to chance level, thereby minimizing between-participant variance and, as a consequence, task reliability.
 As a second example, participants completing an experiment online from their homes may experience more distraction than if they participated in the lab.
-This may in turn lead to an increase in measurement error and a concomitant decrease in reliability.
+This may increase the measurement error, leading to a concomitant decrease in reliability.
 Experimenters therefore cannot assume the reliability of a task measure is constant.
-Ideally then, experimenters should investigate and report the reliability of task measures as part of any analysis or, at the very least, after having made changes to a task or scoring procedure or when administering the task to new sample populations or in new testing contexts.
+Ideally researchers would investigate and report the reliability of task measures as part of any analysis or, at the very least, after having made changes to a task or scoring procedure or when administering the task to new sample populations or in new testing contexts.
 
@@
 
-Though verifying the reliability of cognitive task measures is paramount to successful individual differences research, the reliability of task measures are historically seldom reported \cite{Green2016-xw, Parsons2019-jw}.
-When reported, it has frequently been observed that task measures exhibit lower reliability than what is conventionally considered acceptable for individual differences research (e.g.
-$\rho$ = 0.7 - 0.8) and lower than what is regularly achieved by self-report measures.
-Indeed many studies have now found that task measures to exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}.
-One possible explanation for this finding is the so-called ``reliability paradox'' of cognitive tasks \cite{hedge2018reliability}, which states that the often lackluster reliability of tasks is a result of a mismatch in goals between experimental and individual differences psychological research.
+Although verifying the reliability of cognitive task measures is paramount to individual-differences research, the reliability of task measures is seldom reported \cite{Green2016-xw, Parsons2019-jw}.
+When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered acceptable for individual-differences research ($\rho_{tt}$ on the order of 0.7 - 0.8) and the reliability regularly achieved by self-report measures.
+Indeed, many studies have now found that task measures exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}.
+
+One possible explanation for this finding is the so-called ``reliability paradox'' of cognitive tasks \cite{Hedge2018-lf}, which states that the often lackluster reliability of tasks is a result of a mismatch in goals between experimental and individual-differences psychological research.
 In experimental psychology, the goal is often to demonstrate the existence of a behavioral effect.
-One sure-fire means of increasing the power to detect an effect is to minimize between-participants variance of a task measure.
-This of course is the exact opposite of what is desirable for individual differences research, where between-participants variance is essential to achieving reliable task measures.
-Thus, the tendency in biological psychiatry to adopt the most prominent tasks in experimental psychology --- the ones that most reliably demonstrate a behavioral effect --- may actually hamstring our efforts to study individual differences.
+One means of increasing the power to detect an effect is to minimize between-participants variance.
+This is the exact opposite of what is desirable for individual differences research, where between-participants variance is essential to achieving reliable task measures.
+Thus, the tendency in biological psychiatry to adopt the most prominent tasks in experimental psychology---the ones that most reliably demonstrate a behavioral effect---may actually hamstring our efforts to study individual differences.
 
@@
 
-Regardless we do not believe that task measures are inherently less reliable and that pessimism about task-based individual differences research is warranted.
+Regardless, we do not believe that task measures are inherently less reliable than self-report measures, and that pessimism about task-based individual-differences research is warranted.
 It is possible to (re)design tasks to achieve good reliability, even to the high levels dictated by conventional standards \cite{waltmann2022sufficient, sullivan2022enhancing, kucina2022solution, snijder2022psychometric}.
-The purpose of the current article is to provide a narrative review of approaches to improve task measure reliability.
+The purpose of the current article is to provide a narrative review of approaches to improve task-measure reliability.
 The article is divided into two main parts.
@@
 Then, we introduce a taxonomy of approaches for improving task reliability using concrete examples from the published literature.
-We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual differences research.
+We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual-differences research.
 
@@
 
-We begin with a brief discussion of calculating reliability for cognitive task measures.
-There are two measures of reliability of relevant for task measures: internal consistency and test-retest reliability.
-Internal consistency is the reliability of a measure in a single administration.
-In turn, test-retest reliability (also known as temporal stability) is reliability of a task measure across two or more administrations.
-The two types of reliability provide complimentary information: internal consistency quantifies the consistency of individual difference on a task measure in a given testing session, whereas test-retest reliability quantifies the stability of individual differences across time.
-Even for experimenters primarily interested in cross-sectional experiments, test-retest reliability is a useful index.
-This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant state-dependent factors (e.g.
+There are two measures of reliability of relevance for task measures: internal consistency and test-retest reliability.
+Internal consistency is the reliability of a measure in a single administration of a task.
+Test-retest reliability (also known as temporal stability) is the reliability of a task measure across two or more administrations.
+Even for cross-sectional experiments, where stability over time is of no scientific interest, test-retest reliability is a useful index.
+This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant, state-dependent factors (e.g.
 current mood, fatigue).
-Given sufficient time between testing sessions, test-retest reliability should be less biased state factors.
-A second point to consider with respect to these two indices is that, whereas high internal consistency is virtually always desirable for individual difference studies, high test-retest reliability may not be.
-Test-retest reliability should be interpreted with respect to the construct of interest.
-For example, low 1-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g.
-mood effects on cognition), but may be problematic if the effect is hypothesized to be stable across time (e.g.
-personality effects on cognition).
+Given sufficient time between testing sessions, test-retest reliability should be less biased by state factors, and thus a better estimate of the true cross-sectional task-measure reliability.
+However, whereas high internal consistency is always desirable for individual difference studies, high test-retest reliability may not be, depending on the construct a researcher is intending to measure.
+Low one-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g., mood-dependent attentional biases), but is a problem if individual differences in a cognitive process are hypothesized to be stable across time (e.g., extraversion and social reward processing).
 
@@
 An alternative approach is to calculate the intraclass correlation coefficient (ICC), which decomposes a task measure into true score variance and error variance.
-There are many formulas for calculating ICC \cite{Koo2016-fz}, with the critical distinction being reliability is based on the \textit{consistency} or \textit{absolutely agreement} of a task measure across two administration.
-Consistency-based ICCs are affected only by the relative ordering of participants across time; that is, they are insensitive to systematic changes to the actual values of a task measure across time (e.g.
-practice effects).
-In contrast, absolute agreement-based ICCs measure the degree to which scores are identical across time.
-They are sensitive to changes in the overall mean of a task measure across sessions.
-The type of ICC to use depends on the experimenter's goals.
-If an experimenter cares only about the rank-ordering of participants across time, a consistency-based ICC (or Pearson correlation) is preferable.
+There are many formulas for calculating ICC \cite{Koo2016-fz}, with the critical distinction being whether reliability is based on the \textit{consistency} or \textit{absolutely agreement} of a task measure across two administrations.
+Consistency-based ICCs are affected only by the relative ordering of participants across time; that is, they are insensitive to systematic changes to the actual values of a task measure across time (e.g., due to practice effects on task performance).
+In contrast, absolute-agreement based ICCs measure the degree to which scores are stable across time.
+The type of ICC to use depends on the experimenter's goals and the ultimate use of the task measure.
+% desiderata from the measured individual differences.
+%If an experimenter cares only about the rank-ordering of participants across time, a consistency-based ICC (or Pearson correlation) is preferable.
 Conversely, if systematic differences in a task measure across task administrations is a concern, then an absolute agreement-based ICC should be preferred.
 
-Calculating the internal consistency is more complicated for task measures.
-The most common measure of internal consistency, Cronbach's $\alpha$, requires certain assumptions that are unrealistic for task data (e.g.
-equivalence of trials, uncorrelated measurement error).
-As such, internal consistency for task measures is instead usually calculated via split-$K$ reliability instead, where $K$ indicates the number of splits.
-Split-half reliability ($K=2$) is most frequently chosen.
+Calculating the internal consistency of a task measure is more complicated.
+The most common measure of internal consistency, Cronbach's $\alpha$, requires certain assumptions that are unrealistic for many task data (e.g., equivalence of trials, uncorrelated measurement error).
+As such, internal consistency for task measures is instead usually calculated via split-$K$ reliability, where $K$ indicates the number of splits.
+Split-half reliability ($K=2$) is most common.
 A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be biased when partitions are not equally good measures of a behavior \cite{Green2016-xw, pronk2022can}.
-For example, first-second splitting (i.e.
-partitioning the data by the first and second halves of an experiment) may be confounded by practice, fatigue, or other linear time effects.
-Under such conditions, first-second partitioning will spuriously decrease the similarity of data across partitions, thereby increasing estimates of measurement noise and underestimating reliability.
-In contrast, odd-even splitting (i.e.
-partitioning the data by odd and even trials) may be confounded when behavior across trials is non-independent (i.e.
-correlated errors).
+For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to differences between the partitions due to practice, fatigue, or other linear time effects.
+In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may be confounded when behavior across trials is non-independent (i.e., measurement error is correlated across trials).
 Under such conditions, odd-even splitting will artificially inflate the similarity of data across partitions, thereby decreasing estimates of measurement noise and overestimating reliability.
-Therefore, where possible, a permutation-based approach to calculating split-half reliability has been recommended \cite{pronk2022can, Parsons2019-jw}.
-Briefly, the idea is to compute the average reliability across many thousands of random partitions of the data into halves.
-For task measures derived from cognitive models, it may be unreasonably computationally intensive to employ such an approach.
-In which case, it may instead be preferable to design tasks with at least two independent blocks where the model is fit to each block independently and reliability is calculated using the model parameters estimated from each.
+Therefore, where possible, a permutation-based approach to calculating split-half reliability is recommended \cite{pronk2022can, Parsons2019-jw}.
+Here, reliability is averaged across many thousands of random partitions of the data into halves.
+For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the task measure for each new subset of the data.
+Moreover, for learning tasks that are commonly used in computational psychiatry, cognitive model-based task measures cannot be estimated from only a subset of the trials due to inherent non-independence of task behavior across trials.
+In these cases, one can design tasks with at least two independent blocks.
+Then a model is fit to each block independently and reliability is calculated using the model parameters estimated from each.
 
-A final point is that, when calculating reliability for task measures, traditional sum or mean score estimates of performance (e.g.
-proportion correct responses, mean response time) may substantially underestimate reliability \cite{Rouder2019-am, haines2020learning}.
-These summary scores of task performance neglect trial-by-trial variability in performance, which can result in noisier estimates of task performance and consequently diminished estimates of reliability.
-Instead, it is often preferable to use trial-level hierarchical (mixed effects) models that can simultaneously estimate between- and within-participant (i.e.
-trial-level) variability in behavior.
-By directly modeling trial-by-trial variability, estimates of participants' performance are effectively denoised.
-Furthermore, the use of hierarchical models allows for partial pooling of trial data across participants, which further increases the precision of estimates of participants' performance.
-Together these properties can substantially boost estimates of reliability \cite{Rouder2019-am, haines2020learning}.
-In practice, multiple studies have demonstrated how the use of hierarchical trial-level models directly improves estimates of task measure reliability \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}.
-For a detailed discussion of hierarchical models in the context of task reliability, see Haines (this issue).
+As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability in cases where some response variability across trials can be accounted for rather than treated as noise \cite{Rouder2019-am, haines2020learning}.
+These summary scores are noisy (and thus less reliable) estimates of task performance as they neglect trial-by-trial variability.
+Instead, it is often preferable to use generative or statistical models that account for trial-level variability in behavior.
+By directly modeling trial-by-trial variability, estimates of participants' performance are effectively de-noised.
+It is also advisable to use hierarchical models that partially pool data across participants, which in turn further increases the precision of estimates of participants' performance.
+Together, these methods can substantially boost estimates of reliability \cite{Rouder2019-am, haines2020learning}, as multiple studies have demonstrated  \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}.
+For a detailed discussion of generative models in the context of task reliability, see Haines (this issue).
 
@@
 
-As defined above, the reliability of a task measure is proportion of variance attributable to between-participant differences relative to measurement error.
-Thus, there are two major strategies for improving the reliability of a measure: by increasing between-participant variability or by decreasing measurement error.
-In what follows, we discuss approaches for accomplishing both strategies in turn.
-Where appropriate, we highlight studies from the published literature that are exemplary for imporiving the reliability of a task measure by implementing a particular strategy.
+As defined above, the reliability of a task measure is the proportion of variance attributable to between-participant differences relative to measurement error.
+Thus, the two major strategies for improving the reliability of a measure are to increase between-participant variability or decrease measurement error.
+In what follows, we discuss approaches for accomplishing each objective in turn.
+Where appropriate, we highlight studies that are exemplary for improving the reliability of a task measure by implementing a particular strategy.
 
@@
 Siegelman and colleagues \cite{siegelman2017measuring} noted the consequences of floor effects on reliability in the context of statistical learning tasks.
-In statistical learning tasks, participants must learn to identify subtle patterns in the transition probabilities underlying a continuous sequence of stimuli.
+In such tasks, participants must learn to identify subtle patterns in the transition probabilities underlying a continuous sequence of stimuli.
 In reanalyzing archival datasets, Siegelman and colleagues found a majority of participants were at chance-level performance in discriminating between legitimate and foil sequence patterns; consequently, the reliability of conventional proportion correct measures suffered.
-In response, the authors designed a new version of a statistical learning task involving stimulus sequences that ranged more widely in their difficulty to learn.
-Only a minority of participants were at chance-level performance on this new task and, as such, the reliability of proportion correct scores improved.
-Similarly, in developing an abbreviated working memory task, Oswald and colleagues \cite{oswald2015development} found that they could remove the easiest trials --- those with ceiling level performance --- with virtually no change to task reliability.
-This is because working memory trials with performance at ceiling are incapable of differentiating ability across participants and therefore cannot contribute significantly to the reliability of the task.
+In response, the authors designed a new statistical learning task involving stimulus sequences that ranged more widely in their difficulty to learn.
+Only a minority of participants showed chance-level performance on this new task and, as such, the reliability of proportion correct scores improved.
+Similarly, in developing an abbreviated working-memory task, Oswald and colleagues \cite{oswald2015development} found that they could remove the easiest trials---those with ceiling level performance---with virtually no change to task reliability.
+This is because those trials are incapable of differentiating ability across participants and therefore cannot contribute significantly to the reliability of the task.
 
-Experimenters who are administering a task to a new population should be wary of range restriction effects.
-Cognitive tasks calibrated for one group of participants may be too easy or difficult for a different group.
+Researchers administering a task to a new population should be especially wary of range-restriction effects.
+Cognitive tasks calibrated for one group of participants may not be adequately sensitive for others due to being too easy or difficult for a different group.
 For example, Arnon and colleagues \cite{arnon2020current} found that statistical learning tasks developed for adults were too difficult for young children and therefore yielded unreliable discrimination scores in that population.
-Similarly, Kyllonen and colleagues \cite{kyllonen2019general} were motivated to develop a battery of fluid reasoning measures for highly educated adults due to observed ceiling effects in performance when using preexisting fluid reasoning tasks in this population.
-In sum, experimenters should be cautious when administering preexisting tasks to new populations; the tasks calibrated for one group may not be adequately sensitive for others.
+Similarly, Kyllonen and colleagues \cite{kyllonen2019general} were motivated to develop a battery of fluid-reasoning measures for highly educated adults due to observed ceiling effects in performance when using preexisting fluid-reasoning tasks in this population.
 
@@
 
-A related issue for task reliability is practice effects, or when participants' performance on a task improves with repeated administrations.
+A related issue for task reliability is practice effects, where participants' performance on a task improves with repeated administrations.
 Practice effects are relatively common for cognitive tasks \cite{hausknecht2007retesting, scharfen2018retest}.
@@
 One strategy for minimizing practice effects is simply to increase the time interval between task administrations.
-The more time that elapses between sessions, the greater the probability that participants have forgotten task-specific knowledge or strategies \cite{hausknecht2007retesting, scharfen2018retest}.
-Of course, this solution may not always be possible or desirable, especially if an experimenter is only able to or specifically interested in studying a behavior over a short time period.
+The more time that elapses between sessions, the greater the probability that participants will have forgotten task-specific knowledge or strategies \cite{hausknecht2007retesting, scharfen2018retest}.
+Of course, this solution may not always be possible or desirable, especially if a researcher is only able to or specifically interested in studying a behavior over a short time period.
+Moreover, some forms of learning do not easily dissipate with time.
 
@@
 In the beads task, participants are presented with two jars containing beads of two colors in equal but opposite ratios.
-In one trial of the task, a predetermined sequence of beads is drawn from one jar.
-After participants must decide which jar beads were being drawn from or request to see more beads.
+In each trial of the task, a predetermined sequence of beads is drawn from one jar.
+Participants must decide which jar beads were being drawn from or request to see more beads.
 In a typical version of the task, the same sequence of bead draws is used across all trials.
 McLean and colleagues found that participants were aware that the sequence repeated across trials and, as a consequence, became more erratic in their decision to witness more bead draws with additional trials.
-In response, the authors developed a new version of the beads task that included distractor sequences of bead draws.
-This new design prevent participants from becoming aware of the target sequence, which in turn motivated participants to respond more consistently thereby improving the reliability of participants' information seeking scores.
+In response, the authors developed a new version of the task that included distractor sequences of bead draws.
+This new design was effective in preventing participants from becoming aware of the target sequence, which in turn resulted in more consistent responding, which improved the reliability of participants' information seeking scores.
 
@@
 
-The preceding sections described potential threats to between-participants variability, but did not address how to actually increase it.
-A primary strategy for increasing between-participants variability (and therefore reliability) is to enhance the experimental effect.
-Amplifying the magnitude of an experimental effect (e.g.
-making a task more challenging, increasing the potency of a stress or affective induction) typically increases the range of participants' responses to it.
-Two recent studies highlight this approach in the context of cognitive control tasks.
+The preceding sections described potential threats to between-participants variability, but not approaches to improve it.
+A primary strategy for increasing between-participants variability is to enhance the experimental effect.
+Amplifying the magnitude of an experimental effect (e.g., making a task more challenging, increasing the potency of affect induction) typically increases the range of participants' responses to it.
+For example, Kucina and colleagues \cite{kucina2022solution} investigated the reliability of cognitive conflict effects in new versions of several standard cognitive-control tasks (e.g., Stroop, Flanker, Simon) that amplified cognitive interference via two task design features.
+First, they combined multiple sources of cognitive interference in the same task, for example by combining the Stroop and Simon effects to create a ``Stroopon'' task.
+Second, for a subset of trials, they required participants to make multiple responses based on both relevant and irrelevant stimuli attributes.
+Compared to the standard version of these tasks, these experimental manipulations had the effect of increasing task demands, which resulted in greater between-participants variance, and higher reliability as a consequence.
+(See also Snijder and colleagues \cite{snijder2022psychometric} for a similar redesign of classic cognitive-control tasks that improved reliability by increasing proactive control demands.)
 
-Kucina and colleagues \cite{kucina2022solution} investigated the reliability of cognitive conflict effects in new versions of several standard cognitive control tasks (e.g.
-Stroop, Flanker, Simon) that amplified cognitive interference via two task design features.
-First, they combined multiple sources of cognitive interference in the same task (e.g.
-combining the Stroop and Simon effects to create a ``Stroopon'' task).
-Second, for a subset of trials, they required participants to make multiple responses based on both relevant and irrelevant stimuli attributes.
-Compared to the standard version of these tasks, these experimental manipulations had the effect of increasing task demands, which resulted in greater between-participants variance and reliability as a consequence.
-Similarly, Snijder and colleagues \cite{snijder2022psychometric} studied the reliability of cognitive conflict effects in multiple redesigned cognitive control tasks (e.g.
-Stroop, AX-CPT, Cued Task Switching, Sternberg) that required of participants either more proactive control (i.e.
-maintained vigilance) or reactive control.
-The authors found that the proactive control versions of the task were more demanding, which similarly led to larger cognitive interference effects and ultimately task reliability.
+% Similarly, Snijder and colleagues \cite{snijder2022psychometric} studied the reliability of cognitive conflict effects in multiple redesigned cognitive-control tasks (e.g., Stroop, AX-CPT, Cued Task Switching, Sternberg) that required of participants either more proactive control (i.e., maintained vigilance) or more reactive control than the traditional versions of these tasks.
+The authors found that the proactive control versions of the task were more demanding, which similarly led to larger cognitive interference effects and better task reliability.
 
-For decision-making tasks, or other tasks requiring participants to choose between two or more discrete options, a related strategy is to calibrate the difficulty of trials to the average ability of the population of interest.
+A related strategy is to calibrate the difficulty of the task to the average ability of the population of interest.
 For example, consider a task trial with two response options.
-Given the Bernoulli distribution, the variance in responses on this trial is maximized when the probability of choosing either response is equal (i.e.
-50\%/50\%).
-Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items are around the average ability of the sample \cite{gulliksen1945relation, ferrando2007external} (or somewhere above this point when guessing the correct response is an issue \cite{lord1952relation, feldt1993relationship}).
-Of course, this design principle is only helpful to the degree that an experimenter  precisely knows the average ability level of their participants.
+Given the Bernoulli distribution, the variance in responses on this trial will be maximal when the probability of choosing either response is equal.
+Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items is matched to the average ability of the sample \cite{gulliksen1945relation, ferrando2007external} (or slightly higher if participants can guess the correct response \cite{lord1952relation, feldt1993relationship}).
+Of course, this design principle is only helpful to the degree that a researcher knows the average ability level of their participants.
 If this is unknown or poorly characterized, then it is instead desirable to design a task with trials spanning a uniform range of difficulty.
@@
 While convenient, undergraduate students from a single university are likely to be relatively homogeneous in their cognitive profiles.
-It may be worthwhile instead to recruit participants from the community or from an online labor market (e.g.
-Amazon Mechanical Turk, Prolific Academic, CloudResearch Panels).
-With regard to the latter, it is worth mentioning that, because online participants are typically completing experiments from their homes, they may be more likely to be distracted or multi-tasking during an experiment \cite{newman2021data}.
+It may be worthwhile instead to recruit participants from the community or from an online labor market (e.g., Amazon Mechanical Turk, Prolific Academic, CloudResearch Panels).
+With regard to the latter, because online participants typically complete experiments from their homes or other poorly-controlled environments, they are more likely to be distracted or to multi-task during an experiment \cite{newman2021data}.
 Thus, when recruiting online samples, experimenters should take special care to ensure that an increase in between-participants variance is not offset by a concomitant increase in measurement noise.
@@
 In generalizability theory, the relationship between reliability and the number of trials defined as:
-
 \begin{equation}
-    \rho = \frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E / n}
+    \rho = \frac{\sigma^2_T}{\sigma^2_T + \frac{\sigma^2_E}{n}}
 \end{equation}
-
 where $\sigma^2_T$ is the true between-participants variance, $\sigma^2_E$ is measurement error (i.e.
@@
 In practice, this relationship often holds (e.g.
-\cite{paap2016role, cooper2017role}) but there are sometimes exceptions \cite{price2015empirical, klingelhoefer2022robust}.
-Increasing the number of task trials only benefits reliability if measurement error is random.
-If increasing task length results in participant fatigue or boredom, then measurement noise may systematically increase and, as a consequence, reliability will suffer.
-Furthermore, even if increasing the trial number does not induce fatigue, it may still be impractical for other reasons.
-Because increasing the trial number yields diminishing marginal improvements for reliability, achieving a desired level of reliability may require prohibitively long experiments, especially when the task is completed as part of a larger battery.
+\cite{paap2016role, cooper2017role}) with some exceptions \cite{price2015empirical, klingelhoefer2022robust}.
+Notably, increasing the number of task trials only benefits reliability if measurement error is random.
+If increasing task length results in participant fatigue or boredom, measurement noise may systematically increase and reliability will suffer.
+Increasing the number of trials may be impractical for other reasons, and because of diminishing marginal improvements for reliability, achieving a desired level of reliability through this means alone may require prohibitively long experiments.
 
@@
 The reliability of a task measure can be improved by including in an experiment only the most discriminating stimuli.
-For example, in the context of an emotion recognition task, stimuli of good discriminability with would be those where participants with good emotion recognition ability consistently correctly identify the displayed emotion and were participants with poor ability consistently incorrectly identify the displayed emotion \cite{keutmann2015generating}.
-In contrast, stimuli with poor discriminability --- those for which performance between high and low ability participants is indistinguishable --- will lead to more measurement noise and decreased reliability.
-In experiments where stimuli are intended to be unique and distinguishable, it is important to ensure that participants are not confusing the identities of the stimuli.
-Improving both the linguistic and visual distinctness of stimuli may prevent this sort of confusion and therefore aid reliability \cite{yoo2022importance}.
+For example, in the context of an emotion recognition task, stimuli of good discriminability would be those where participants with good emotion-recognition ability consistently correctly identify the displayed emotion while participants with poor ability consistently incorrectly identify the displayed emotion \cite{keutmann2015generating}.
+In contrast, stimuli with poor discriminability --- those for which performance between high- and low-ability participants is indistinguishable --- will lead to more measurement noise and decreased reliability.
+In experiments where stimuli are intended to be unique and distinguishable, improving both the linguistic and visual distinctness of stimuli may prevent confusion and therefore aid reliability \cite{yoo2022importance}.
 
-Other design features of an experiment may affect reliability.
+Other design features of an experiment that are specific to the task mechanics may affect reliability.
 Consider for example the dot-probe task, in which participants must disengage attention from a distracting image on one part of the screen in order to identify and respond to the orientation of a pair of dots elsewhere on the screen.
-Previous research has found that dot-bottom trials, where a participant disengage from a distracting stimulus located at the top of the screen and then saccade to the bottom of the screen, are more reliable than dot-top trials \cite{price2015empirical, aday2019extended}.
-An explanation for this finding is that, if participants tend to have a gaze bias towards the top of the screen, a saccade away from the top of the screen might require a stronger level of disengagement.
-As such, dot-bottom trials are sometime found to be a more robust signal of attentional bias.
-In summary, the actual mechanics of a task trial may have profound influence on task reliability
+Dot-bottom trials, where a participant must disengage from a distracting stimulus located at the top of the screen and saccade to the bottom of the screen, are more reliable than dot-top trials \cite{price2015empirical, aday2019extended}.
+This has been explained by suggesting that participants are biased to gaze towards the top of the screen, thus saccading away from the top of the screen requires a stronger level of disengagement, allowing such trials to better measure attentional bias.
 
-It is worth stressing that clear, concise instructions are essential for task reliability.
-When participants are unsure of what they are intended to do in an experiment, their behavior is likely to be more erratic and random, thereby diminishing reliability.
-(Clear instructions may also work to ensure the validity of an experiment by discouraging participants from using cognitive strategies not of interest to the experimenter.) In their 10 simple rules paper for designing cognitive experiments, Barbosa and colleagues \cite{barbosa2022practical} provide practical suggestions for writing task instructions.
+It is worth stressing that clear instructions are essential for task reliability.
+When participants are unsure of what they are intended to do in an experiment, their behavior is likely to be more variable across time (as their understanding of the task evolves) and across participants (due to different interpretation of instructions), thereby diminishing reliability.
+(Clear instructions may also work to ensure the validity of an experiment by discouraging participants from using strategies not of interest to the experimenter.) In their ``10 simple rules'' paper for designing cognitive experiments \cite{barbosa2022practical}, Barbosa and colleagues provide practical suggestions for writing task instructions.
 When adapting a task for a new population, experimenters should ensure that the instructions are still appropriate.
@@
 
-Another strategy to reduce measurement error is to incorporate practice trials into an experiment.
-Practice trials can help to minimize the effects of nuisance factors such as performance anxiety or unfamiliarity with the response modality.
-Moreover, practice trials give participants participants an additional opportunity to make sure they understand the task instructions.
-Thus, practice trials can help participants reach a more consistent ``steady state'' of behavior, thereby reducing the noisiness of their responses and increasing reliability \cite{alexander2003effects}.
-If experimenters are unable to include a standalone practice block in their experimenters due to time constraints, they can implicitly incorporate practice trials by designating and discarding (or modeling separately) the first handful of trials during an experiment \cite{mclean2018towards}.
+Another strategy to reduce measurement error is to make use of a practice phase.
+Practice trials can help to minimize the effects of nuisance factors such as performance anxiety or unfamiliarity with the response modality, and give participants an opportunity to make sure they understand the task instructions.
+Thus, practice trials can help participants reach a ``steady state'' of responding, thereby reducing the noisiness of their responses across task trials and increasing reliability \cite{alexander2003effects}.
+Practice can take the form of a standalone practice block or by designating as such and discarding (or modeling separately) the first few trials of an experiment \cite{mclean2018towards}, though the latter may not fully allow participants to explore response options during the practice.
 
 Yet another strategy to diminish measurement error is to ``gamify'' an experiment.
-The gamification of cognitive tasks, --- incorporating (video) game design elements into cognitive tasks --- can increase participant engagement and motivation during an experiment \cite{sailer2017gamification}.
-Keeping participants motivated and engaged may counter the would-be effects of boredom and fatigue on task reliability.
-For example, Kucina and colleagues \cite{kucina2022solution} cite task gamification as an important factor that contributed to the improved the reliability of their cognitive control tasks.
+Incorporating (video) game design elements into cognitive tasks can increase participant engagement and motivation \cite{sailer2017gamification}, countering the would-be effects of boredom and fatigue on task reliability.
+For example, Kucina and colleagues \cite{kucina2022solution} cite task gamification as an important factor that contributed to the improved the reliability of their cognitive-control tasks.
 Similarly, Verdejo and colleagues \cite{verdejo2021unified} partially attribute the improved reliability of their impulsivity task battery to gamified task design.
@@
 
-When the parameters from cognitive models are used as summaries of participants' task performance, another means to improve measure reliability is to decrease estimation noise.
-The estimation noise of a parameter given an experiment and model can be quantified through simulation studies \cite{wilson2019ten, palminteri2017importance}.
-In a simulation study, an experimenter generates artificial data for the experiment using representative model parameters.
-The experimenter attempts to recover the model parameters by fitting the model to the simulated data.
+When parameters from cognitive models are used as indices of participants' task performance, another means to improve measure reliability is to decrease estimation noise.
+The estimation noise of a parameter given an experiment and model can be quantified through simulation \cite{wilson2019ten, palminteri2017importance}.
+Here, an experimenter generates artificial data for the experiment using representative model parameters and then attempts to recover the model parameters by fitting the model to the simulated data.
 Estimation noise is the inverse of the (relative or absolute) agreement between the true and recovered parameters.
 Alterations to experimental design can improve parameter recovery and estimation noise, and multiple frameworks have been proposed for testing and improving experimental designs to aid parameter recovery \cite{broomell2014parameter, melinscak2020computational}.
-Parameter recovery can also be affected by model estimation method \cite{Lerche2017-yg, waltmann2022sufficient}.
-As mentioned above, the partial pooling properties of hierarchical Bayesian models can be especially beneficial for improving parameter recovery and decreasing estimation noise \cite{katahira2016hierarchical}.
+Parameter recovery can also be affected by the model estimation method \cite{Lerche2017-yg, waltmann2022sufficient}.
+In particular, the partial pooling properties of hierarchical Bayesian models can be especially beneficial for improving parameter recovery and decreasing estimation noise \cite{katahira2016hierarchical}.
 
-A related approach is to use adaptive experimental designs \cite{myung2013tutorial}.
-In an adaptive experiment, the design of an experiment is generated on-line in order to present each participant with stimuli or trial types that are matched to their particular response patterns or ability levels.
+A related approach is to use adaptive experimental designs \cite{myung2013tutorial}, where the trials of an experiment are designed in real time so as to present each participant with stimuli or trial types that are matched to their particular response patterns or ability levels.
 Though undoubtedly a more complex experimental design, adaptive experiments have the advantage of selecting the most informative trials for resolving the ability or preference level of a participant (e.g.
 as measured by a cognitive model).
-Furthermore, adaptive designs  may require fewer trials to achieve precision measurement, which can save experimenters (and participants) time and money.
-Adaptive designs have been successfully translated in cognitive research for example in working memory \cite{gonthier2018measuring} and delay discounting \cite{ahn2020rapid}.
-For a detailed discussion of adaptive design experiments, see Ahn (this issue).
+Adaptive designs have been successfully used in cognitive research, for example to study working memory \cite{gonthier2018measuring} and delay discounting \cite{ahn2020rapid}.
+For a detailed discussion of adaptive-design experiments, see Ahn (this issue).
 
 Parameter estimation may be further improved by leveraging additional information.
-For example, latent variables may be more accurately measured through the inclusion of demographic variables or other covariates during estimation \cite{anders2018improved, Curran2016-bt}.
-The basic idea is that, if there are any associations between model parameters and covariates, they can aid in resolving parameter estimates.
+For example, latent variables may be more accurately measured through the inclusion of demographic variables or other covariates \cite{Curran2016-bt} that, if associated to model parameters, can aid in resolving parameter estimates.
 An extension of this idea is to utilize joint modeling of dependent variables; that is, to design models where multiple observed trial-level variables are predicted simultaneously.
 For example, the joint modeling of choice and response time has been found to improve the precision and reliability of estimated parameters in cognitive ability testing \cite{bertling2018using} and reinforcement learning \cite{ballard2019joint, shahar2019improving}.
-It is also possible to incorporate physiological and/or neural correlates of behavior, such as skin conductance response, fMRI BOLD signal, and EEG.
+It is also possible to incorporate physiological and/or neural correlates of behavior, such as skin conductance response, fMRI BOLD signal, and EEG \cite{palestro2018tutorial}.
 
@@
 
-In the context of reliability, difference scores deserve special treatment.
-Difference scores describe a measure where a measure of a participant's performance from one condition is subtracted from their performance in another.
-An example of a difference score is the classic Stroop interference effect; that is, the difference in average reaction time between congruent and incongruent trials.
-The value of difference scores is straightforward: difference scores allow experimenters to isolate particular cognitive processes (e.g.
-processing cognitive conflict) and to control for irrelevant sources of variance (e.g.
-perceptual processing, motor ability).
-Given their utility, difference scores are ubiquitous in experimental psychology.
-
-The challenge for difference scores is that their reliability is a function of the reliability of their components \textit{and} the correlation between the components:
-
+Difference scores deserve special treatment in the context of reliability.
+Difference scores subtract a measure of a participant's performance in one condition from their performance in another.
+An example is the classic Stroop interference effect: the average reaction-time difference between congruent and incongruent trials.
+Difference scores are commonly used because they allow experimenters to isolate particular cognitive processes (e.g.
+processing cognitive conflict) while controlling for other sources of variance (e.g.
+perceptual processing, motor ability) through the subtraction of conditions that share that variance.
+The challenge is that the reliability of a difference score is a function of the reliability of each of its components and the correlation between the components:
 \begin{equation}
@@
 \end{equation}
-
-where $\sigma^2_x$ and $\sigma^2_y$ are the variances of task measures \textit{x} and \textit{y}; $\rho_{xx'}$ + $\rho_{yy'}$ are the reliabilities of task measures \textit{x} and \textit{y}; and $\rho_{xy}$ is the correlation between task measures \textit{x} and \textit{y} \cite{trafimow2015defense, chiou1996reliability}.
+where $\sigma^2_x$ and $\sigma^2_y$ are the variances of task measures \textit{x} and \textit{y}, $\rho_{xx'}$ and $\rho_{yy'}$ are the reliabilities of task measures \textit{x} and \textit{y}, and $\rho_{xy}$ is the correlation between task measures \textit{x} and \textit{y} \cite{chiou1996reliability}.
 When the variances of the two measures are equal, this reduces to:
-
 \begin{equation}
@@
 \end{equation}
-
-From the above equation, it is readily apparent that the reliability of a difference score measure is diminished to the extent that its components are correlated.
-In experimental psychology, two measures derived from the same task will often be correlated due to shared domain-general cognitive processes.
-Thus, it will often (if not always) be the case that difference scores derived from task measures will be less reliable than their components.
-This presents a challenge for designing reliable difference score measures.
-
-\subsubsection{Identify alternative measures}
-
-An alternative approach to improving the reliability of difference scores is simply to avoid using them in the first place.
-That is, experimenters may opt to design tasks or analyses in such a way to avoid the necessity of difference scores in the first place.
-This recommendation has a long history in experimental psychology.
-Indeed, because difference scores will virtually always be less reliable, many authors have advocated to simply abandon them \cite{lord1956measurement, cronbach1970we, edwards2001ten}.
-
-What then are the alternatives to difference scores?
-Draheim and colleagues \cite{draheim2019reaction} provide an in-depth review of alternatives to difference scores in the context of response time measures, though much of their discussion is applicable to difference scores in general.
-One possibility is to simply use the component measures (e.g.
-performance on incongruent trials in a Stroop task alone).
-Of course, because component scores will be contaminated other sources of variance, such as baseline performance, interpreting component scores should be done with caution.
-Another approach is to identify alternative measures of task performance.
-For example, intra-individual response time variability has been identified as a correlate in executive control, with aberrations in psychopathology broadly \cite{kofler2013reaction}, that can be measured reliably \cite{saville2011stability}.
+From the above equation, one can see that the reliability of a difference score measure is diminished to the extent that its components are correlated.
+Two measures derived from the same task will often be correlated due to shared domain-general cognitive processes.
+Thus, it will often (if not always) be the case that difference scores derived from task measures will be less reliable than the average reliability of their components.
 
@@
 
-When difference scores are unavoidable, there still steps experimenters can take to improve their reliability.
-As the equations above suggest, the reliability of difference scores can be improved in three ways: (1) improve the reliability of the component measures, (2) increase the differential variability of the task measures, and (3) minimize the correlation between the task measures.
-The first strategy has been the focus of the discussion so far.
+The equations above suggest three steps experimenters can take to improve the reliability of difference scores: (1) improve the reliability of the component measures, (2) increase the difference in the magnitude of the variances of the two task measures, and (3) minimize the correlation between the task measures.
+The first strategy has been our focus so far.
 The second approach deserves further comment.
-As noted elsewhere \cite{trafimow2015defense, chiou1996reliability}, the reliability of a difference score measure increases as ratio of measure variances increases.
-Figure XX shows that, even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the deviation ratio is high.
-This speaks to another advantage of increasing the between-participants variability of a measure (insofar that an experimenter increases one, but not the other, measure).
-This result might help explain the improved reliability of the difference score measures studied by Kucina and colleagues \cite{kucina2022solution}.
+As discussed elsewhere \cite{chiou1996reliability}, the reliability of a difference score measure increases as the difference between (or the ratio of) the variances of the component measures increases.
+The intuitive explanation for why this is the case is that, as the variance of one (but not the other) component grows, so to does the proportion of unshared reliable variance.
+Figure \ref{fig:fig03} shows that even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the variance ratio (i.e.
+$\sigma_x / \sigma_y$) is high.
+This speaks to another advantage of increasing the between-participants variability of a measure (insofar that there is an increase in one, but not the other, measure).
+This fact may help explain the improved reliability of the difference score measures studied by Kucina and colleagues \cite{kucina2022solution}.
+
+\begin{figure}[tp]
+    \centering
+    \includegraphics[width=1\textwidth]{figures/fig02.png}
+    \caption{Difference score reliability as a function of the variances and reliabilities of its component measures.
+(A) When the variances of the component measures are equal ($\sigma_x / \sigma_y = 1$), larger correlations between the measures substantially diminish the reliability of a difference score measure.
+(B) When the variances of the component measures are unequal (e.g., here, $\sigma_x / \sigma_y = 4$), even large correlations between the measures impact the reliability of a difference score measure only minimally.
+Here, the reliabilities of the component measures are assumed to be equal.}
+    \label{fig:fig03}
+\end{figure}
 
@@
 individual differences in speed-accuracy preferences; \cite{draheim2019reaction}).
-Noting this, Rey-Mermet and colleagues designed a number of ``response-deadline'' executive control tasks, where participants had a limited amount of time to respond during a trial.
-Specifically, the duration of the response deadline was calibrated for each participant individually in order to achieve a fixed accuracy level in blocks of neutral trials, which was then used as the deadline for both congruent and incongruent trials.
-They reasoned the calibration procedure had several important advantages.
-First, it controls for individual differences in processing speed that contribute to performance in both congruent and incongruent trials.
-Second, it also controls for individual differences in strategy, because, regardless of whether a participant was biased towards speed or accuracy, inefficient executive control would result in lower accuracy.
-They found that, with the calibration procedure, the reliability of an accuracy difference score (incongruent - congruent) was as good or better than what had previously been reported for executive control tasks.
-Thus, controlling for shared variance across measures --- that is, purifying measures --- can help to improve task reliability.
+Rey-Mermet and colleagues designed a number of ``response-deadline'' executive control tasks where participants had a limited amount of time to respond during a trial.
+The duration of the response deadline was calibrated for each participant individually such that they achieve a fixed accuracy level in blocks of neutral trials, and  then used as the deadline for both congruent and incongruent trials.
+The calibration procedure controls for individual differences in processing speed that contribute to performance in both congruent and incongruent trials.
+It also controls for individual differences in strategy, as regardless of whether a participant was biased towards speed or accuracy, inefficient executive control would result in lower accuracy.
+With this calibration procedure, the researchers found that the reliability of an accuracy difference score (incongruent - congruent) was as good or better than what had previously been reported for executive control tasks.
+Thus, controlling for shared variance across measures---that is, purifying measures---can help to improve task reliability.
+
+\subsubsection{Identify alternative measures}
+
+Rather than improving the reliability of difference scores, one can simply avoid using them in the first place.
+This recommendation has a long history in experimental psychology.
+Indeed, because difference scores will virtually always be less reliable, many authors have advocated to abandon them \cite{lord1956measurement, cronbach1970we, edwards2001ten}.
+
+What then are the alternatives to difference scores?
+Draheim and colleagues \cite{draheim2019reaction} provide an in-depth review of alternatives to difference scores in the context of response-time measures, though much of their discussion is applicable to difference scores in general.
+One possibility is to simply use the component measures (e.g.
+performance on incongruent trials in a Stroop task alone).
+Of course, because component scores will be contaminated other sources of variance, such as baseline performance, interpreting component scores should be done with caution.
+Another approach is to identify alternative measures of task performance.
+For example, intra-individual response time variability has been identified as a correlate of executive control that can be measured reliably \cite{saville2011stability} and is altered in psychopathology \cite{kofler2013reaction}.
 
 \section{Discussion}
+
+% outline
+% 1) summarize
+% 2) validity: three examples, (1) cognitive control + convergent validity, (2) interpretability of task parameters, (3) predictive validity of tasks (impulsivity), (4) ecological validity.
+maybe even start with rouder paper: some tasks may not be able to made reliable!
+other challenges of scaling up, repeated testing, naturalistic contexts (cite germine)
+% 3) conclusion: pessimism not warranted. requirements: measuring reliability, sharing case studies for improvement, sharing task code (jspsych-demos)
+
+% inspiration 
+% The findings presented here convey the importance of scrutinizing the psychometric properties of cognitive tasks, especially those intended for use in individual differences studies.
+Reliability is not always reported or even examined by researchers, and the present results also demonstrate why variability and reliability should be considered and reported in the context of one another.
+Future research efforts related to individual differences in cognition could benefit from incorporating psychometric techniques.
+This may hopefully lead to improvements in the design and evaluation of cognitive tasks for the purposes of individual differences research.
 

### 2022-10-26 manuscript/main.tex
checkpoint progress

@@
 % set document class
-\documentclass[a4paper,notitlepage,12pt]{article}
+\documentclass[a4paper,12pt]{article}
 
@@
 % specify the title
-\title{Improving the reliability of cognitive task measures: A narrative review}
+\title{Improving the reliability of cognitive task measures: \break A narrative review}
 
@@
 
+\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena.
+However, many recent studies have found that task measures exhibit poor reliability, hampering their utility for individual differences research.
+The purpose of this article is to provide a narrative review of approaches to improve the reliability of cognitive task measures.
+First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks.
+Then, we introduce a taxonomy of approaches for improving task reliability.
+Where appropriate, we highlight studies that are exemplary for improving the reliability of specific task measures.
+We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual-differences research.}
+
 \section{Introduction}
@@
 Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry.
-Variability in performance on cognitive tasks provides experimenters with crucial insights into the relationship between individual differences in cognitive processing and behavioral phenomena outside of the laboratory or clinic (e.g.
+Individual differences in performance on cognitive tasks provide researchers with crucial insights into the relationship between cognitive processing and behavioral phenomena in the clinic and in our everyday lives (e.g.
 reward sensitivity in a monetary-incentivized task and anhedonia in depression).
@@
 This is a necessary property of tasks to be used for studying individual differences; where there is no variation in performance, there are no individual differences to study.
-The validity of a task measure concerns whether it actually measures what it purports to measure.
+The validity of a task measure concerns whether it actually measures what it intends to measure.
 Finally, the reliability of a task measure characterizes the degree to which it consistently rank-orders participants across measurement instances.
 Put another way, a task measure is reliable if, assuming participants have not changed, it produces the same score for each participant over time.
-The focus of this review is task reliability, which is a prerequisite for validity.
-
+The focus of this review is task reliability.
 % P3: definition and Importance of reliability
@@
 
-In classical test theory, the variance in observed scores on a task measure $X$ is the sum of the true score variance $T$, reflecting real individual differences in the latent construct of interest, and measurement error $E$, i.e., $X = T + E$.
-The reliability of a measure is defined as the proportion of variance attributable to the true score variance relative to total variance: $\rho = \frac{T}{T + E}$.
+In classical test theory, the variance in observed scores on a task measure $x$ is the sum of the true score variance $\sigma^2_T$, reflecting real individual differences in the latent construct of interest, and measurement error $\sigma^2_E$, i.e., $\sigma^2_x = \sigma^2_T + \sigma^2_E$.
+The reliability of a measure is defined as the proportion of variance attributable to the true score variance relative to total variance: $\rho_{xx'} = \frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$.
 Thus, reliability quantifies the magnitude of individual differences relative to the noisiness of a task measure; the larger the reliability of a task measure, the more it reflects true individual differences rather than reflecting noise.
@@
 As the reliability of two measures decreases, the number of participants required to detect an association increases.
-Even with a large sample size of $N=500$, a measure with reliability $\rho_{tt} = 0.5$ will only allow detection of correlations above $0.3$, which are uncharacteristically high in the field.}
+Even with a large sample size of $N=500$, two measures with moderate reliability ($\rho_{xx'} = \rho_{yy'} = 0.5$) will only reliably detect true correlations above $0.3$, which are ostensibly high for the field.}
     \label{fig:fig01}
@@
 To further complicate matters, the reliability of a task measure is not absolute -- it reflects interactions between task design, the participants, and the context in which the task is administered.
-Indeed, task reliability can vary as a function of parameters of the experiment such as the specific stimulus set, number of trials, time limits \cite{paap2016role, cooper2017role}); sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in clinic, online); response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); and scoring or estimation method \cite{Rouder2019-am, haines2020learning}.
+Indeed, task reliability can vary as a function of parameters of the experiment such as the specific stimulus set, number of trials, time limits \cite{paap2016role, cooper2017role}; sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in clinic, online); response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); and scoring or estimation method \cite{Rouder2019-am, haines2020learning}.
 For example, a cognitive task originally designed for use with an adult population may prove too difficult for children.
@@
 As a second example, participants completing an experiment online from their homes may experience more distraction than if they participated in the lab.
-This may increase the measurement error, leading to a concomitant decrease in reliability.
+This may increase measurement error, leading to a concomitant decrease in reliability.
 Experimenters therefore cannot assume the reliability of a task measure is constant.
-Ideally researchers would investigate and report the reliability of task measures as part of any analysis or, at the very least, after having made changes to a task or scoring procedure or when administering the task to new sample populations or in new testing contexts.
+Ideally researchers would investigate and report the reliability of task measures as part of any individual differences research or, at the very least, after having made changes to a task or scoring procedure or when administering the task to new sample populations or in new testing contexts.
 
@@
 
-Although verifying the reliability of cognitive task measures is paramount to individual-differences research, the reliability of task measures is seldom reported \cite{Green2016-xw, Parsons2019-jw}.
-When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered acceptable for individual-differences research ($\rho_{tt}$ on the order of 0.7 - 0.8) and the reliability regularly achieved by self-report measures.
+Although verifying the reliability of cognitive task measures is paramount to individual differences research, the reliability of task measures is seldom reported \cite{Green2016-xw, Parsons2019-jw}.
+When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered the minimum acceptable level for individual-differences research ($\rho_{tt}$ on the order of 0.7 -- 0.8) and the reliability regularly achieved by self-report measures.
 Indeed, many studies have now found that task measures exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}.
@@
 
-As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability in cases where some response variability across trials can be accounted for rather than treated as noise \cite{Rouder2019-am, haines2020learning}.
-These summary scores are noisy (and thus less reliable) estimates of task performance as they neglect trial-by-trial variability.
+As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning}.
+These summary scores are contaminated by trial-level noise that, in the absence of sufficiently large number of trials, increase the measurement error in (and thus diminish the reliability of) these measures of task performance.
 Instead, it is often preferable to use generative or statistical models that account for trial-level variability in behavior.
@@
 perceptual processing, motor ability) through the subtraction of conditions that share that variance.
-The challenge is that the reliability of a difference score is a function of the reliability of each of its components and the correlation between the components:
+The challenge is that the reliability of a difference score is a function of the reliability of each of its components \textit{and} the correlation between the components:
 \begin{equation}
@@
 
-The equations above suggest three steps experimenters can take to improve the reliability of difference scores: (1) improve the reliability of the component measures, (2) increase the difference in the magnitude of the variances of the two task measures, and (3) minimize the correlation between the task measures.
+The equations above suggest three steps experimenters can take to improve the reliability of difference scores: (1) improve the reliability of the component measures, (2) increase the relative difference between the variances of two task measures, and (3) minimize the correlation between the task measures.
 The first strategy has been our focus so far.
@@
 As discussed elsewhere \cite{chiou1996reliability}, the reliability of a difference score measure increases as the difference between (or the ratio of) the variances of the component measures increases.
-The intuitive explanation for why this is the case is that, as the variance of one (but not the other) component grows, so to does the proportion of unshared reliable variance.
-Figure \ref{fig:fig03} shows that even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the variance ratio (i.e.
-$\sigma_x / \sigma_y$) is high.
-This speaks to another advantage of increasing the between-participants variability of a measure (insofar that there is an increase in one, but not the other, measure).
-This fact may help explain the improved reliability of the difference score measures studied by Kucina and colleagues \cite{kucina2022solution}.
+Intuitively this is because, as the variance of one (but not the other) component grows, so too does the proportion of unshared reliable variance.
+Figure \ref{fig:fig03}A shows that even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the ratio of the variances (i.e.
+$\sigma_x / \sigma_y$) is sufficiently greater or less than 1.
+(Also worth noting is that component variances essentially function as weights as in a weighted average, such that difference score reliability more reflects the reliability of the component measure with the larger variance; Figure \ref{fig:fig03}B.) This speaks to another advantage of increasing the between-participants variability of a measure of performance in one experiment condition insofar that there is not an accompanying increase in perormance variability in a second condition.
+This may explain why making more difficult the incongruent trials in the tasks designed by Kucina and colleagues improved the reliability of their difference score measures \cite{kucina2022solution}.
 
@@
     \includegraphics[width=1\textwidth]{figures/fig02.png}
-    \caption{Difference score reliability as a function of the variances and reliabilities of its component measures.
-(A) When the variances of the component measures are equal ($\sigma_x / \sigma_y = 1$), larger correlations between the measures substantially diminish the reliability of a difference score measure.
-(B) When the variances of the component measures are unequal (e.g., here, $\sigma_x / \sigma_y = 4$), even large correlations between the measures impact the reliability of a difference score measure only minimally.
-Here, the reliabilities of the component measures are assumed to be equal.}
+    \caption{\small Difference score reliability as a function of the variances and reliabilities of its component measures.
+(A) Difference score reliabilities as the ratio of the component measures' variances ($\sigma_x / \sigma_y$) increase, but the component reliabilities are held fixed ($\rho_{xx'} = \rho_{yy'} = 0.8$).
+When the variances of the component measures are equal ($\sigma_x / \sigma_y = 1$), larger correlations between the measures substantially diminish the reliability of a difference score measure.
+When the variances are unequal ($\sigma_x / \sigma_y > 1$), even large correlations between the measures are less deleterious for reliability.
+(B) Difference score reliabilities as the component reliabilities change, but the ratio of the component measures' variances is fixed ($\sigma_x / \sigma_y = 2$).
+When the variances are unequal, the reliability of a difference score reflects more the component measure with the larger variance.}
     \label{fig:fig03}
@@
 
-\section{Discussion}
+\section{Conclusion}
 
-% outline
-% 1) summarize
-% 2) validity: three examples, (1) cognitive control + convergent validity, (2) interpretability of task parameters, (3) predictive validity of tasks (impulsivity), (4) ecological validity.
-maybe even start with rouder paper: some tasks may not be able to made reliable!
-other challenges of scaling up, repeated testing, naturalistic contexts (cite germine)
-% 3) conclusion: pessimism not warranted. requirements: measuring reliability, sharing case studies for improvement, sharing task code (jspsych-demos)
+In this brief article, we have reviewed issues and research regarding cognitive task-measure reliability.
+Specifically, we introduced a taxonomy of approaches for improving cognitive task-measure reliability, highlighting studies that are exemplary for improving the reliability of specific task measures.
+We hope we have made clear the importance of calculating (and reporting) the reliability of task measures intended for use in psychiatric research.
+We also hope we have provided a useful guide for experimenters who wish to design a new task, or improve an existing task, to study individual differences in cognitive processing.
 
-% inspiration 
-% The findings presented here convey the importance of scrutinizing the psychometric properties of cognitive tasks, especially those intended for use in individual differences studies.
-Reliability is not always reported or even examined by researchers, and the present results also demonstrate why variability and reliability should be considered and reported in the context of one another.
-Future research efforts related to individual differences in cognition could benefit from incorporating psychometric techniques.
-This may hopefully lead to improvements in the design and evaluation of cognitive tasks for the purposes of individual differences research.
+Before concluding, we must emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures.
+Task measures may be reliable but show poor convergent validity \cite{snijder2022psychometric, eckstein2022interpretation}, raising questions about whether we actually measuring the constructs we intended to measure.
+Similarly, task measures may be reliable but exhibit poor ecological validity \cite{Steiner2021-oq}, thus being poor proxies for cognition in real-world settings.
+Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables we care about (e.g.
+symptoms, treatment response).
+Finally, there are many other practical considerations (e.g.
+duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully to the clinic or naturalistic contexts at scale \cite{germine2021toward}.
+
+Despite the challenges of making task-measures reliable and valid, we remain optimistic about their current and future use in biological psychiatry.
+We firmly believe that, with further research efforts towards developing, documenting, and sharing reliable task paradigms, our field can make great strides towards understanding, predicting, and ultimately relieving psychiatric illness.
+
+\section{Code availability}
+
+The code used to generate the figures in this manuscript are publicly available at \break \url{https://github.com/nivlab/biopsych-reliability-review}.
+
+\section{Acknowledgments}
+
+This research was funded in part by a National Science Foundation Graduate Research Fellowship.
+
+\section{Disclosures}
+
+The authors have no conflicts of interest to declare.
 

### 2022-10-28 manuscript/main.tex
latest (final?) draft

@@
 \setlength{\affilsep}{2em}   % set the space between author and affiliation
-\author[1]{Samuel Zorowitz}
+\author[1,*]{Samuel Zorowitz}
 \author[1,2]{Yael Niv}
@@
 \affil[2]{Department of Psychology, Princeton University, USA}
+\affil[*]{Corresponding author {\small (zorowitz@princeton.edu)}}
 
@@
 
-\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena.
-However, many recent studies have found that task measures exhibit poor reliability, hampering their utility for individual differences research.
-The purpose of this article is to provide a narrative review of approaches to improve the reliability of cognitive task measures.
+\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena across individuals.
+However, many recent studies have found that task measures exhibit poor reliability, which hampers their utility for individual-differences research.
+Here we provide a narrative review of approaches to improve the reliability of cognitive task measures.
 First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks.
@@
 Where appropriate, we highlight studies that are exemplary for improving the reliability of specific task measures.
-We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual-differences research.}
+We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing one, to achieve sufficient reliability for use in individual-differences research.}
 
@@
 Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry.
-Individual differences in performance on cognitive tasks provide researchers with crucial insights into the relationship between cognitive processing and behavioral phenomena in the clinic and in our everyday lives (e.g.
+Individual differences in performance on cognitive tasks provide researchers with crucial insights into the relationship between cognitive processing and behavioral phenomena in the clinic and in everyday life (e.g.
 reward sensitivity in a monetary-incentivized task and anhedonia in depression).
@@
 As the reliability of two measures decreases, the number of participants required to detect an association increases.
-Even with a large sample size of $N=500$, two measures with moderate reliability ($\rho_{xx'} = \rho_{yy'} = 0.5$) will only reliably detect true correlations above $0.3$, which are ostensibly high for the field.}
+Even with a large sample size of $N=500$, two measures with moderate reliability ($\rho_{xx'} = \rho_{yy'} = 0.5$) will only reliably detect true correlations above $0.3$, which are likely high for individual-differences cognitive research \cite{owens2021recalibrating}.}
     \label{fig:fig01}
@@
 Although verifying the reliability of cognitive task measures is paramount to individual differences research, the reliability of task measures is seldom reported \cite{Green2016-xw, Parsons2019-jw}.
-When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered the minimum acceptable level for individual-differences research ($\rho_{tt}$ on the order of 0.7 -- 0.8) and the reliability regularly achieved by self-report measures.
+When they are reported, task measures frequently exhibit lower reliability than what is conventionally considered the minimum acceptable level for individual-differences research ($\rho_{xx'}$ on the order of 0.7 -- 0.8) and the reliability regularly achieved by self-report measures.
 Indeed, many studies have now found that task measures exhibit moderate-to-low reliability \cite{Hedge2018-lf, Frey2017-uz, Enkavi2019-oh, Von_Bastian2020-tm, Nitsch2022-pe, verdejo2021unified}.
@@
 However, whereas high internal consistency is always desirable for individual difference studies, high test-retest reliability may not be, depending on the construct a researcher is intending to measure.
-Low one-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g., mood-dependent attentional biases), but is a problem if individual differences in a cognitive process are hypothesized to be stable across time (e.g., extraversion and social reward processing).
+For example, low one-month test-retest reliability may not be problematic for an index of a transient cognitive process (e.g., mood-dependent attentional biases), but is a problem if individual differences in a cognitive process are hypothesized to be stable across time (e.g., extraversion and social reward processing).
 
@@
 Consistency-based ICCs are affected only by the relative ordering of participants across time; that is, they are insensitive to systematic changes to the actual values of a task measure across time (e.g., due to practice effects on task performance).
-In contrast, absolute-agreement based ICCs measure the degree to which scores are stable across time.
+In contrast, absolute-agreement-based ICCs measure the degree to which scores are stable across time.
 The type of ICC to use depends on the experimenter's goals and the ultimate use of the task measure.
-% desiderata from the measured individual differences.
-%If an experimenter cares only about the rank-ordering of participants across time, a consistency-based ICC (or Pearson correlation) is preferable.
-Conversely, if systematic differences in a task measure across task administrations is a concern, then an absolute agreement-based ICC should be preferred.
 
@@
 A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be biased when partitions are not equally good measures of a behavior \cite{Green2016-xw, pronk2022can}.
-For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to differences between the partitions due to practice, fatigue, or other linear time effects.
+For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to differences between the partitions resulting from practice, fatigue, or other linear time effects.
 In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may be confounded when behavior across trials is non-independent (i.e., measurement error is correlated across trials).
@@
 Here, reliability is averaged across many thousands of random partitions of the data into halves.
-For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the task measure for each new subset of the data.
-Moreover, for learning tasks that are commonly used in computational psychiatry, cognitive model-based task measures cannot be estimated from only a subset of the trials due to inherent non-independence of task behavior across trials.
+For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the model for each new subset of the data.
+Moreover, for learning tasks that are commonly used in computational psychiatry, cognitive-model based task measures cannot be estimated from only a subset of the trials due to inherent non-independence of task behavior across trials.
 In these cases, one can design tasks with at least two independent blocks.
-Then a model is fit to each block independently and reliability is calculated using the model parameters estimated from each.
+The model can then be fit to each block independently and reliability calculated using the model parameters estimated from each.
 
 As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning}.
-These summary scores are contaminated by trial-level noise that, in the absence of sufficiently large number of trials, increase the measurement error in (and thus diminish the reliability of) these measures of task performance.
+This is because such summary scores are contaminated by trial-level noise that, in the absence of sufficiently large number of trials, increases the measurement error (and thus diminishes reliability).
 Instead, it is often preferable to use generative or statistical models that account for trial-level variability in behavior.
@@
 Of course, this solution may not always be possible or desirable, especially if a researcher is only able to or specifically interested in studying a behavior over a short time period.
-Moreover, some forms of learning do not easily dissipate with time.
+Moreover, some forms of learning do not easily dissipate with time \cite{schiller2010preventing}.
 
-A second strategy is to design tasks so as to prevent or discourage the formation of task-specific strategies.
+A second strategy is to use a combination of clear instructions and practice trials to help participants reach stable performance from the start of an experiment (see the ``Improving experiment designs'' section below).
+Another strategy is to design tasks so as to prevent or discourage the formation of task-specific strategies.
 For example, McLean and colleagues \cite{mclean2018towards} investigated the repeatability of the beads task.
@@
 
-% Similarly, Snijder and colleagues \cite{snijder2022psychometric} studied the reliability of cognitive conflict effects in multiple redesigned cognitive-control tasks (e.g., Stroop, AX-CPT, Cued Task Switching, Sternberg) that required of participants either more proactive control (i.e., maintained vigilance) or more reactive control than the traditional versions of these tasks.
-The authors found that the proactive control versions of the task were more demanding, which similarly led to larger cognitive interference effects and better task reliability.
-
 A related strategy is to calibrate the difficulty of the task to the average ability of the population of interest.
@@
 Given the Bernoulli distribution, the variance in responses on this trial will be maximal when the probability of choosing either response is equal.
-Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items is matched to the average ability of the sample \cite{gulliksen1945relation, ferrando2007external} (or slightly higher if participants can guess the correct response \cite{lord1952relation, feldt1993relationship}).
+Thus, aggregating across many trials, between-participants variance is maximized --- and task reliability is improved --- when the difficulty of all items is matched to the average ability of the sample \cite{gulliksen1945relation} (or slightly higher if participants can guess the correct response \cite{lord1952relation, feldt1993relationship}).
 Of course, this design principle is only helpful to the degree that a researcher knows the average ability level of their participants.
@@
 trial-level variance), and $n$ is the number of trials.
-In practice, this relationship often holds (e.g.
-\cite{paap2016role, cooper2017role}) with some exceptions \cite{price2015empirical, klingelhoefer2022robust}.
+In practice, this relationship often holds \cite{paap2016role, cooper2017role} though with some exceptions \cite{price2015empirical, klingelhoefer2022robust}.
 Notably, increasing the number of task trials only benefits reliability if measurement error is random.
@@
 When participants are unsure of what they are intended to do in an experiment, their behavior is likely to be more variable across time (as their understanding of the task evolves) and across participants (due to different interpretation of instructions), thereby diminishing reliability.
+Clear instructions thus help to ensure that participants show stable behavior from the start of the experiment.
 (Clear instructions may also work to ensure the validity of an experiment by discouraging participants from using strategies not of interest to the experimenter.) In their ``10 simple rules'' paper for designing cognitive experiments \cite{barbosa2022practical}, Barbosa and colleagues provide practical suggestions for writing task instructions.
@@
 The second approach deserves further comment.
-As discussed elsewhere \cite{chiou1996reliability}, the reliability of a difference score measure increases as the difference between (or the ratio of) the variances of the component measures increases.
-Intuitively this is because, as the variance of one (but not the other) component grows, so too does the proportion of unshared reliable variance.
+As discussed elsewhere \cite{chiou1996reliability}, the reliability of a difference score measure increases as the difference (or the ratio) between the variances of the component measures increases.
+Intuitively, this is because as the variance of one (but not the other) component grows, so too does the proportion of unshared reliable variance.
 Figure \ref{fig:fig03}A shows that even when the correlation of two measures is large, it is possible to achieve acceptable reliability insofar as the ratio of the variances (i.e.
-$\sigma_x / \sigma_y$) is sufficiently greater or less than 1.
-(Also worth noting is that component variances essentially function as weights as in a weighted average, such that difference score reliability more reflects the reliability of the component measure with the larger variance; Figure \ref{fig:fig03}B.) This speaks to another advantage of increasing the between-participants variability of a measure of performance in one experiment condition insofar that there is not an accompanying increase in perormance variability in a second condition.
-This may explain why making more difficult the incongruent trials in the tasks designed by Kucina and colleagues improved the reliability of their difference score measures \cite{kucina2022solution}.
+$\sigma_x / \sigma_y$) is sufficiently different from 1.
+(It is also worth noting that component variances essentially function as weights in determining the overall reliability, such that the difference-score reliability reflects more the reliability of the component measure with the larger variance; Figure \ref{fig:fig03}B.) This speaks to the advantage of increasing the between-participants variability of a measure of performance in one experiment condition without increasing performance variability in a second condition.
+This may explain why Kuchina and colleagues improved the reliability of their difference score by making only the incongruent trials more difficult \cite{kucina2022solution}.
 
-\begin{figure}[tp]
+\begin{figure}[t]
     \centering
@@
     \caption{\small Difference score reliability as a function of the variances and reliabilities of its component measures.
-(A) Difference score reliabilities as the ratio of the component measures' variances ($\sigma_x / \sigma_y$) increase, but the component reliabilities are held fixed ($\rho_{xx'} = \rho_{yy'} = 0.8$).
-When the variances of the component measures are equal ($\sigma_x / \sigma_y = 1$), larger correlations between the measures substantially diminish the reliability of a difference score measure.
+(A) Difference-score reliability as the ratio of the component measure variances ($\sigma_x / \sigma_y$) increases, with component reliabilities held fixed ($\rho_{xx'} = \rho_{yy'} = 0.8$).
+When component measures have equal variances ($\sigma_x / \sigma_y = 1$), large correlations between the measures substantially diminish the reliability of a difference score measure.
 When the variances are unequal ($\sigma_x / \sigma_y > 1$), even large correlations between the measures are less deleterious for reliability.
-(B) Difference score reliabilities as the component reliabilities change, but the ratio of the component measures' variances is fixed ($\sigma_x / \sigma_y = 2$).
-When the variances are unequal, the reliability of a difference score reflects more the component measure with the larger variance.}
+(B) Difference score reliability as the component reliabilities change, with the ratio of the component-measure variances fixed ($\sigma_x / \sigma_y = 2$).
+When the variances are unequal, the reliability of a difference score reflects more the component measure with the larger variance (here, component $x$).}
     \label{fig:fig03}
@@
 
-In this brief article, we have reviewed issues and research regarding cognitive task-measure reliability.
-Specifically, we introduced a taxonomy of approaches for improving cognitive task-measure reliability, highlighting studies that are exemplary for improving the reliability of specific task measures.
+We have briefly reviewed issues and research regarding the reliability of cognitive-task measures.
+Specifically, we introduced a taxonomy of approaches for improving cognitive task-measure reliability, highlighting example studies that improved the reliability of specific task measures.
 We hope we have made clear the importance of calculating (and reporting) the reliability of task measures intended for use in psychiatric research.
-We also hope we have provided a useful guide for experimenters who wish to design a new task, or improve an existing task, to study individual differences in cognitive processing.
+We also hope we have provided a useful guide for experimenters who wish to design a new task, or to improve an existing task, in order to study individual differences in cognitive processing.
 
-Before concluding, we must emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures.
-Task measures may be reliable but show poor convergent validity \cite{snijder2022psychometric, eckstein2022interpretation}, raising questions about whether we actually measuring the constructs we intended to measure.
+We emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures.
+Task measures may be reliable but show poor convergent validity \cite{snijder2022psychometric, eckstein2022interpretation}, raising questions about whether they actually measure the constructs they are intended to measure.
 Similarly, task measures may be reliable but exhibit poor ecological validity \cite{Steiner2021-oq}, thus being poor proxies for cognition in real-world settings.
-Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables we care about (e.g.
+Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables of interest (e.g.
 symptoms, treatment response).
 Finally, there are many other practical considerations (e.g.
-duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully to the clinic or naturalistic contexts at scale \cite{germine2021toward}.
+task duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully in the clinic or in naturalistic environments at scale \cite{germine2021toward}.
 
-Despite the challenges of making task-measures reliable and valid, we remain optimistic about their current and future use in biological psychiatry.
+Despite the challenges of making task-measures reliable and valid, we are optimistic about their current and future use in biological psychiatry.
 We firmly believe that, with further research efforts towards developing, documenting, and sharing reliable task paradigms, our field can make great strides towards understanding, predicting, and ultimately relieving psychiatric illness.
@@
 
-The code used to generate the figures in this manuscript are publicly available at \break \url{https://github.com/nivlab/biopsych-reliability-review}.
+The code used to generate the figures in this manuscript is publicly available at \break \url{https://github.com/nivlab/biopsych-reliability-review}.
 
@@
 
-This research was funded in part by a National Science Foundation Graduate Research Fellowship.
+We thank Rachel Bedder for helpful feedback on this manuscript.
+This project was made possible with support from the National Center for Advancing Translational Sciences (NCATS), a component of the National Institute of Health (NIH), under award number UL1TR003017a and a National Science Foundation Graduate Research Fellowship.
 
@@
 
+\section{Citation diversity statement}
+
+Recent work in several fields of science has identified a bias in citation practices such that papers from women and other minority scholars are under-cited relative to the number of such papers in the field \cite{dworkin2020extent, bertolero2020racial}.
+Here we sought to proactively consider choosing references that reflect the diversity of the field in thought, form of contribution, gender, race, ethnicity, and other factors.
+First, we obtained the predicted gender of the first and last author of each reference by using databases that store the probability of a first name being carried by a woman \cite{dworkin2020extent}.
+By this measure (and excluding self-citations to the first and last authors of our current paper), our references contain 8.92\% woman(first)/woman(last), 19.37\% man/woman, 21.92\% woman/man, and 49.79\% man/man.
+This method is limited in that a) names, pronouns, and social media profiles used to construct the databases may not, in every case, be indicative of gender identity and b) it cannot account for intersex, non-binary, or transgender people.
+Second, we obtained predicted racial/ethnic category of the first and last author of each reference by databases that store the probability of a first and last name being carried by an author of color \cite{ambekar2009name, sood2018predicting}.
+By this measure (and excluding self-citations), our references contain 4.81\% author of color (first)/author of color(last), 15.59\% white author/author of color, 16.92\% author of color/white author, and 62.68\% white author/white author.
+This method is limited in that a) names and Florida Voter Data to make the predictions may not be indicative of racial/ethnic identity, and b) it cannot account for Indigenous and mixed-race authors, or those who may face differential biases due to the ambiguous racialization or ethnicization of their names.
+We look forward to future work that could help us to better understand how to support equitable practices in science.
+
 \printbibliography

### 2023-01-09 manuscript/letter.tex
manuscript revisions + letter to reviewers

% set document class
\documentclass[a4paper,12pt]{article}

% load packages
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage{hyperref}
\usepackage{csquotes}

% specify bibliography source
\addbibresource{letter.bib}
\renewcommand*{\bibfont}{\small}

% specify link style
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=blue
}

% specify some custom commands
\renewcommand{\baselinestretch}{1.2}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand*{\Affilfont}{\normalsize}

% set up the author block
\setlength{\affilsep}{2em}   % set the space between author and affiliation
\author[1]{Samuel Zorowitz}
\author[1,2]{Yael Niv}
\affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
\affil[2]{Department of Psychology, Princeton University, USA}

% specify the title
\title{Response to reviewers for ``Improving the reliability of cognitive task measures: A narrative review''}

% turn date off
\date{}

% Define paragraph formatting
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\begin{document}

\maketitle

\section*{Formatting note}

In the document below, reviewers' comments are indicated in bold typeface. Quoted text from the revised manuscript is presented in quotation blocks; within these excerpts, text that is new is indicated in italicized typeface. All updates to text in the revised manuscript are detailed below.

\section*{Reviewer \#1}

\textbf{This review discusses strategies for improving the reliability of commonly used cognitive tasks, with examples from the literature. Overall I thought this was a nicely written review that makes a constructive contribution to the literature. My comments are minor suggestions where the authors could expand on the critical discussion of some points.}

\textbf{1. It may be worth considering Rouder et al.'s (2019) preprint \cite{rouder2019most}, which discusses some of these same issues (e.g. how far hierarchical models can get us, whether the required trial numbers for some tasks are prohibitively large). It is perhaps also noting (e.g.) cases where hierarchical models did not improve reliability (Whitehead et al., 2020; \cite{whitehead2020reliability}).}

We thank the reviewer for bringing both papers to our attention. We have cited Rouder et al. at multiple points throughout the revised manuscript, including in the section on calculating reliability to note that hierarchical modeling is not a universal panacea. Specifically, the text now reads (beginning on page 5, line 42):

\begin{displayquote}
     Instead, it may be preferable to use trial-level hierarchical models in which observations are organized hierarchically (e.g., individuals within a group, trials within an individual) with variability modeled at both levels. Hierarchical models exert a pooling or regularization effect on person-level variables, in effect correcting for measurement error and improving estimates of reliability [13–15]. The benefits of hierarchical models for estimating reliability has been multiply demonstrated [27, 32–34], \textit{though see [Rouder et al.] for discussion of when these benefits may be limited.}
\end{displayquote}

With respect to Whitehead et al., we would interpret their study as providing further evidence in favor of hierarchical models. Though the reliability of the Simon effect did not differ substantially between hierarchical modeling vs. traditional sum score approaches ($\rho$ = 0.65 vs. $\rho$ = 0.61), the reliability of both the Stroop effect ($\rho$ = 0.65 vs. $\rho$ = 0.50) and Flanker effect ($\rho$ = 0.57 vs. $\rho$ = 0.31) were notably improved.

\textbf{2. Pg. 5 notes that the type of ICC to use depends on the experimenter's goals - it may be worth expanding this point to encompass discussion of other methods. For example, hierarchical models may lead to higher reliability estimates, but this may not be the most informative estimate if subsequent analysis/applications don't have the same statistical properties (e.g. if the measures are used in traditional analyses, or if it is intended to be used as an individual-level diagnostic).}

We thank the reviewer for this point. However, we are unsure of the sort of situation that the reviewer has in mind. For example, if a researcher is using a task measure as an individual-level diagnostic (e.g., for treatment selection for a particular patient), group-level priors from an already-fit hierarchical model could be used to regularize a model fit to a particular patient's data. Because of this, and due to word limit constraints, we have elected to forego adding this point to the main text.

\textbf{3. Pg. 6: ``Practice effects are not inherently an issue for reliability — especially if an experimenter is only interested in the consistency, but not the absolute agreement, of participants' performance over time...''. This appears to assume that practice effects would be constant across participants, i.e. that every participant would learn the task at the same rate, therefore not changing the rank ordering. This may require evidence/consideration. With both practice and fatigue effects, there is also a possibility that task length interacts with participants.}

We thank the reviewer for this important point. We have now clarified the text to acknowledge that practice effects may impact reliability insofar that they are differentially expressed by participants. The text now reads (page 7, lines 12-13):

\begin{displayquote}
    Practice effects are not inherently an issue for reliability — especially if an experimenter is only interested in the consistency, but not the absolute agreement, of participants’ performance over time — but they can become a pernicious issue if \textit{they are exhibited differentially across participants} or if they are severe enough to induce ceiling effects.
\end{displayquote}

\section*{Reviewer \#2}

\textbf{Summary: In the current study, the authors review the literature on the psychometrics of cognitive task measures with a focus on the construct of measurement reliability. They overview how poor reliability has impacted individual difference research using cognitive task measures and then discuss various different perspectives on how to resolve issues associated with poor reliability.}

\textbf{Strengths: I think this is a very well written piece! The authors cover quite a broad range of perspectives in a small amount of space, and they provide good directions for future researcher on reliability and cognitive tasks.}

\textbf{Major Suggestions: I only have a few ``major'' suggestions that I believe help correct, clarify, or strengthen the
authors’ arguments:}

\textbf{1. On page 2, the authors note that ``Reliability is therefore a prerequisite for validity: an unreliable task measure reflects measurement error and not the construct of interest.'' I understand the reliability is often discussed as such a prerequisite, but I think it is a bit of an oversimplification. This often-repeated statement assumes that there is some reliability threshold that, once reached, makes it possible for a measure to be valid. In other words, that one cannot make valid inference when reliability is low. However, I don't believe this is the case. e.g., using equation 1 in the main text, if measures X and Y have a true correlation of yet poor reliabilities of $r_T = 0.8$ yet poor reliabilities of $\rho_X = 0.3$ and $\rho_Y = 0.3$, I can expect to observe a correlation of $r_O= r_T \sqrt{(\rho_X \cdot \rho_Y} = 0.24$. If I assume this observed correlation is the true correlation, then this is an invalid inference. However, if we account for unreliability and disattenuate the observed correlation, we get a valid inference. This is a simple example, but the field of decision theory is full of other counter examples—so long as we account for uncertainty, we can obtain valid inferences (albeit our inferences will be very uncertain if reliability is very low). It seems that the phrase I quoted above takes a black-and-white approach to reliability and validity, assuming a measure is either reliable or not (and subsequently allowed to be valid or not). I think the above points are important because the ``black-and-white view'' of reliability is apparently widely held—I think it is largely responsible for why researchers have been sounding alarm bells re: individual difference research using cognitive tasks being untrustworthy. I wonder if the authors could make it more clear that we can indeed make valid inference/decisions when reliability is low, we just need to account for uncertainty/low reliability when drawing inference (which may in some cases stop us from making a decision dependent on our risk preferences, context, etc.).}

We thank the reviewer for this thoughtful point. We believe, however, the reviewer is conflating valid \textit{inference} and valid \textit{measurement}. We do not dispute that researchers can make valid inferences using invalid measures. To continue the example above: if we disattenuate a correlation between two measures with poor reliability, as Rouder et al. \cite{rouder2019most} show, we can expect large uncertainty around our estimate. Insofar that we clearly reflect that uncertainty (e.g., ``the true correlation may be as small as -0.6 or as large as 0.6''), we have made a valid inference. That says little, however, about the validity of our measurement. If the reliability of our measure is so poor that our measure predominantly reflects something other than the construct we intend to measure (e.g. noise), then it is invalid.

With that said, the reviewer touches upon an important point: the false dichotomy between ``acceptable'' and ``unacceptable'' reliability based on conventions from elsewhere in psychology. We believe the uncritical adoption of standards for reliability (e.g., acceptable reliability $\geq$ 0.8) is to the detriment of biological psychiatry. (Moreover, in our view, this is at least part of the reason why some researchers view task-based individual-difference research as untrustworthy.) As such, we have addressed this issue in a new paragraph in the discussion (page 14, lines 1-14):

\begin{displayquote}
    \textit{We conclude with two important points. First, although we have discussed the importance of task reliability, we have largely avoided the question of when a task measure is ``reliable enough''. Though it is tempting to fall back on conventional cutoffs (e.g., $\rho \geq 0.7$), what constitutes sufficient reliability in actuality will depend on the goal(s) of the researcher. If the goal is to detect a significant individual-differences correlation, such as between a task measure and self-reported symptom measure, then a task measure with ``unacceptable'' reliability by conventional standards may suffice (e.g., if a researcher has the resources to collect a sample large enough to be adequately powered to detect a correlation at the attenuated magnitude). On the other hand, if a researcher intends to estimate an individual-differences correlation with high precision, or use a task measure in a high stakes setting (e.g., treatment selection for an individual patient), then high reliability may be required. We cannot overstate the value of simulation studies (e.g., [24]) for researchers trying to determine what level of reliability is required to meet their goals and risk preferences.}
\end{displayquote}

We hope this new addition addresses, at least in part, the reviewer's concerns.

\textbf{2. On page 5, the authors note that: ``... Cronbach's $\alpha$, requires certain assumptions that are unrealistic for many task data... As such, internal consistency for task measures is instead usually calculated via split-K reliability... Split-half reliability (K = 2) is most common.'' This claim assumes that split-half reliability metrics make more realistic assumptions re: task data than Cronbach’s $\alpha$. However, Cronbach’s $\alpha$ is analytically equivalent to the average of all possible split-half reliabilities. The way that the paragraph is organized almost implies to me that split-half reliability is qualitatively different from $\alpha$. I think this is partly true dependent on how data are split, but it would be useful for authors to explicitly explain the relation between $\alpha$ and split half reliability. Then, they can segue into their later discussion on how different splitting choices can be made to relax the restrictive assumptions of $\alpha$.}

We thank the reviewer for this important point. We agree that the framing of the text suggests that Cronbach's $\alpha$ and split-half reliability are qualitatively distinct measures, when in actuality they belong to the same family of measures \cite{Green2016-xw}. Given word limit constraints and the fact that this topic has been discussed at length elsewhere \cite{Green2016-xw}, we have chosen to not go into detail here, but rather clarify and qualify the text while pointing readers to more detailed treatments on the topic. The next now reads (beginning on page 5, line 9):

\begin{displayquote}
    Calculating the internal consistency of a task measure is more complicated. \textit{The most common measure of internal consistency is Cronbach's $\alpha$, which is a function of the average correlation across all unique pairs of trials. However, Cronbach's $\alpha$ is an accurate measure of reliability only under assumptions that are unrealistic for many tasks (e.g., equivalence of trials, uncorrelated measurement error; [6, 16]). As such, internal consistency for task measures is instead usually calculated via split-half reliability, where reliability is estimated after trial data have been divided into two halves.} A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be also biased if the data partitions violate either of the two above assumptions (for detailed discussion, see [16, 30]). For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to nonequivalence of the two partitions resulting from practice, fatigue, or other linear time effects. In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may cause bias when behavior across trials is non-independent (i.e., measurement error is correlated across trials), artificially inflating the similarity of data across partitions and thereby decreasing estimates of measurement noise and overestimating reliability. Therefore, where possible, a permutation-based approach to calculating split-half reliability is recommended [6, 30]. Here, reliability is averaged across many thousands of random partitions of the data into halves. \textit{(Insofar that Cronbach's $\alpha$ is analytically equivalent to the average of all possible split-half reliability estimates [31], permutation-based split-half reliability provides an approximation to Cronbach's $\alpha$ while avoiding its problematic assumptions.)}
\end{displayquote}

\textbf{3. On page 5, the authors note that: ``traditional sum or mean estimates… may substantially underestimate task reliability... because such summary scores are contaminated by trial-level noise that... increases the measurement error... By directly modeling trial-by-trial variability, estimates of participants' performance are effectively de-noised. It is also possible to use hierarchical models that partially pool data across participants''. I think that there are two lines of reasoning here that are conflated, making the individual points incorrect as stated. Specifically, it is not the modeling of trial-level variability that ``denoises'' the estimates—it is either: (1) the use of a model of person-level behavior that better characterizes variability, or (2) the combination of trial-level modeling with the hierarchical model. In other words, if I were to only fit a normal distribution using MLE to each person’s RT distribution across timepoints and correlate the resulting $\mu$ parameters (as an estimate of reliability), I would not expect that the $\mu$ parameters would have higher reliability than the sample means (despite the model technically estimating trial-level noise). The hierarchical model induces pooling, which results in a higher estimate for reliability. That said, one of the arguments we make in Haines et al. (2020; ref 13) is that more sophisticated models may lead to better estimates of reliability. e.g., use of a shifted lognormal over a normal model may produce estimates with better reliability because they better characterize the full distribution of behavior (even absent the hierarchical model). I think perhaps the different points about a choice of person-level model (e.g., normal, lognormal, etc. on RT distributions) and choice of group-level model (e.g., normal population distribution across parameters of person-level models) are being (unintentionally) conflated in the quoted statement above, making it appear as if the authors are claiming that estimating variability alone will increase reliability.}

We thank the reviewer for this important feedback. We have revised this section of the manuscript to clarify our description of the benefits of hierarchical and generative models (beginning on page 5, line 40):

\begin{displayquote}
    As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability [13–15]. This is because such summary scores are contaminated by trial-level noise that, in the absence of a sufficiently (possibly prohibitively) large number of trials, increases measurement error (and thus diminishes reliability). \textit{Instead, it may be preferable to use trial-level hierarchical models in which observations are organized hierarchically (e.g., individuals within a group, trials within an individual) with variability 5 modeled at both levels. Hierarchical models exert a pooling or regularization effect on person-level variables, in effect correcting for measurement error and improving estimates of reliability [13–15]. The benefits of hierarchical models for estimating reliability has been multiply demonstrated [27, 32–34], though see [24] for discussion of when these benefits may be limited. Using statistical models that more accurately characterize the latent data-generating process (e.g., using the shifted log-normal distribution to model response times) may also improve reliability estimates [14, 35].} For a detailed discussion of hierarchical and generative models in the context of task reliability, see Haines and colleagues (this issue; [14]).
\end{displayquote}

\textbf{Minor Suggestions:}

\textbf{1. On page 7 the authors note that: ``Amplifying the magnitude of an experimental effect... typically increases the range of participants’ responses to it.'' I think it would be worth noting that this runs counter to the ``reliability paradox'' hypothesis—i.e. that making an effect strong for experimental purposes results in some sort of reduction in between-person variance. If it is ``typical'' that amplifying an experimental effect actually increases between-person variability, then how is this reconciled with the reliability paradox?}

We thank the reviewer for this thoughtful point. As we understand it, the crux of the ``reliability paradox'' hypothesis is that the most replicable findings in experimental psychology are not necessarily replicable because their effects are large in absolute terms (e.g., the average cognitive interference effect on response times is only 50 ms \cite{rouder2019most}), but because they are consistently elicited across participants (i.e., low between-participants variability). Amplifying the magnitude of an experimental effect (e.g., increasing both the mean and variance of a cognitive interference effect) may therefore decrease group-averaged effect sizes while increasing reliability. As such, this is not necessarily counter to the reliability paradox hypothesis.

This having been said, we realize readers may read ``effect'' as ``effect size'' in our sentence. We therefore changed the text as below (page 8, lines 3-6):

\begin{displayquote}
    A primary strategy for increasing between-participants variability is to enhance the experimental \textit{manipulation}. Amplifying the strength of an experimental \textit{manipulation} (e.g., making a task more challenging, increasing the potency of affect induction) typically increases the range of participants' responses to it.
\end{displayquote}

\section*{Reviewer \#3}

\textbf{This is an excellent and accessible review/guide covering a variety of approaches for improving the reliability of cognitive tasks. I think this will find a broad audience and could serve as a great resource for both established researchers and students in research methods classes.}

\textbf{My comments are relatively minor:}

\textbf{1. The introduction overstates the current utility of cognitive tasks in relation to understanding/predicting/diagnosing psychiatric symptoms. I realize this is not the focus of the paper, but I think either the authors should briefly make a better case for cognitive tasks as ``invaluable tools'' in psychiatry or the authors should present a more balanced view of the current state of cognitive tasks in relation to psychiatry.}

We thank the reviewer for this important point and agree that our characterization of the value-add of cognitive tasks to biological psychiatry (so far) was overstated. We have tempered our language in the introduction, which now reads (beginning on page 1, line 20):

\begin{displayquote}
\textit{Cognitive tasks hold great promise for biological psychiatry.} When properly designed, such tasks are capable of isolating and measuring specific cognitive processes. Individual differences in performance on cognitive tasks can therefore provide researchers with crucial insights into the cognitive processes underlying psychiatric phenomena. Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement [1] and cognitive decline [2]. \textit{Cognitive tasks, then, have the potential to be invaluable tools for refining our understanding of psychiatric symptoms and syndromes.} For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.
\end{displayquote}

\textbf{2. On page 2, the re-worded definition of reliability as ``a task measure is reliable if, assuming participants have not changed, it produces the same score for each participant over time'' is a bit misleading and does not encapsulate the important distinction between reliability/consistency and agreement that the authors delve into later in the paper. It might be helpful to soften the definition used here so that readers do not understand reliability to mean literally ``same score''.}

We thank the reviewer for this important comment. We have refined our definition of reliability, so that the text now reads (page 2, lines 10-12):

\begin{displayquote}
    Finally, the reliability of a task measure characterizes the degree to which it consistently measures some feature of participants. \textit{That is, a task measure is reliable if, assuming participants have not changed, it produces the same scores, or the same ordering of scores, for participants within a single testing session or across multiple testing sessions.} This review focuses on task-measure reliability.
\end{displayquote}

\textbf{3. As accessibility seems to be a major goal of this review and since the reliability paradox is a central issue in using cognitive tasks for individual differences research, I think it would be helpful to give a concrete example when discussing the reliability paradox on pages 3 and 4.}

We thank the reviewer for this helpful suggestion. We have now included a brief example in the paragraph describing the ``reliability paradox'' (page 4, line 2-4):

\begin{displayquote}
    One possible explanation for this finding is the so-called ``reliability paradox'' of cognitive tasks [17], which states that the often lackluster reliability of tasks is a result of a mismatch in goals between experimental and individual-differences psychological research. In experimental psychology, the goal is often to demonstrate the existence of a behavioral effect. One means of increasing the power to detect an effect is to minimize between-participants variance. This is the exact opposite of what is desirable for individual differences research, where between-participants variance is essential to achieving reliable task measures. \textit{For example, the Stroop effect is one of the most robust effects in experimental psychology; virtually everyone shows a Stroop effect [23]. However, in part due to this fact, between-participants variance on the Stroop effect is often limited [24].} Thus, the tendency in biological psychiatry to adopt the most prominent tasks in experimental psychology---the ones that most reliably demonstrate a behavioral effect---may actually hamstring efforts to study individual differences.
\end{displayquote}

\textbf{4. Throughout the manuscript, it would be helpful to highlight the magnitude of the reliability improvement that was achieved by each of the highlighted methods (increasing trials, using highly distinguishable stimuli, reducing ceiling effects, enhancing experimental effects, etc.)? I am left wondering whether these methods are capable of changing task reliability scores from unacceptable to good, from good to excellent? Can we expect cognitive task reliability to perform as well as that of self-report measures?}

We thank the reviewer for another helpful suggestion. Where we discuss changes to task designs that resulted in an improvement to reliability, we have now provided descriptions of the magnitude of improvement. For example, when describing McLean and colleagues' efforts to improve the beads task, the text now reads (page 7, line 40):

\begin{displayquote}
    This new design was effective in preventing participants from becoming aware of the target sequence, which in turn resulted in more consistent responding, which improved the reliability of participants' information seeking scores \textit{(from $\rho = 0.62$ to $0.84$)}. 
\end{displayquote}

\textbf{5. While the paper covers a variety of approaches, there is a more in-depth focus on reviewing methods for altering the design of experimental tasks on the front-end rather than reviewing analytic strategies that can be employed to improve reliability, and I agree that these strategies should get the greatest emphasis. However, I do think it would be helpful to expand a bit regarding analytic strategies, as follows:}

\textbf{a. the sentence ``It is possible to (re)design tasks to achieve good reliability, even to the high levels dictated by conventional standards [21-24]'' is misleading as most of the referenced papers here did not alter the task design; rather they used more recent analytic strategies to improve reliability. This should be clarified as it is an important take-away for readers to understand that it may be possible to improve reliability in legacy data from traditional tasks by using newer analytic strategies. Same comment in relation to the statement ``We hope that this article can serve as a helpful guide for experimenters who wish to \textit{design a new task, or improve an existing task}, to achieve sufficient reliability for use in individual-differences research''. Improved reliability may potentially be achieved without redesigning the task itself.}

We thank the reviewer for these important points. We agree that citing Waltmann et al. and Sullivan-Toole et al. in this context is misleading, as improvements in task reliability in those studies stem from the use of improved statistical methods and not alterations to task design. We have removed those two citations from the first sentence in question. We also agree with the second point. The text of the second sentence in question now reads (page 4, lines 19-20):

\begin{displayquote}
    We hope that this article can serve as a helpful guide for experimenters designing a new task, improving an existing task, \textit{or refining their scoring methods} to achieve sufficient reliability for use in individual-differences research.
\end{displayquote}

\textbf{b. Finally, either in the final full paragraph of page 5 or in section 3.2.3 `Reducing parameter estimation noise', I think it is worth briefly highlighting that hierarchical models \textit{can} improve reliability but are not sufficient and that lower-level model parameterization is also critical for reliability improvement, as demonstrated in Brown, et al., 2021 and Sullivan-Toole, et al., 2022. Although, I am biased with regards to this last point, so the authors should use their own judgment.}

We thank the reviewer for this suggestion. We have revised to text to better highlight that using more appropriate trial-level models can also improve estimates of reliability. The text now reads (page 6, lines 5-7):

\begin{displayquote}
    \textit{Using statistical models that more accurately characterize the latent data-generating process (e.g., using the shifted log-normal distribution to model response times) may also improve reliability estimates [14, 35].} For a detailed discussion of hierarchical and generative models in the context of task reliability, see Haines and colleagues (this issue; [14]).
\end{displayquote}

\section*{Reviewer \#4}

\textbf{This manuscript reviews empirical findings and recommendations on reliability of cognitive tasks. The topic is timely and the review covers a range of areas. This manuscript will be a good reference and resource for researchers in this field. I particularly appreciated the attention to special considerations for reliability in cognitive tasks that differ from other applications of reliability (e.g., questionnaires). My comments primarily are about clarifying points to ensure they are useful to readers new to this area as well as suggested additions to ensure the review covers all relevant topics. Given the goal of the paper, some of these comments are more picky than most reviews I do to minimize potential confusion for the reader.}

\textbf{Clarifications:}

\textbf{1. In the intro, page 2, starting line 10, the authors define reliability as consistency in rank-ordering across measurement instances. As they discuss later on, however, reliability may not be a measure of rank order and it can measure within-session as well as between-session consistency.}

We thank the reviewer for this crucial point (see also Reviewer 3, point 2). We have refined our definition of reliability, so that the text now reads (page 2, lines 10-12):

\begin{displayquote}
    Finally, the reliability of a task measure characterizes the degree to which it consistently measures some feature of participants. \textit{That is, a task measure is reliable if, assuming participants have not changed, it produces the same scores, or the same ordering of scores, for participants within a single testing session or across multiple testing sessions.} This review focuses on task-measure reliability.
\end{displayquote}

\textbf{2. Page 2, line 5 - discriminating power or discriminatory power?}

We thank the reviewer for spotting this typo, which has now been corrected.

\textbf{3. Page 3, starting on line 54 - as the authors note later, other factors besides low within-person variability also affect reliability. In particular, difference scores are not introduced until the section on improving reliability but could instead be introduced with other factors affecting reliability here.}

We thank the reviewer for this helpful suggestion. We now mention difference scores as part of the list of factors affecting reliability. The text now reads (page 2, lines 38-39):

\begin{displayquote}
    Indeed, task reliability can vary as a function of experiment parameters (specific stimulus set, number of trials, time limits [8, 9]); sample populations (healthy adults, children, psychiatric patients [9, 10]); testing locations (in clinic, online); response modality (desktop, smartphone, virtual reality [11, 12]); \textit{scoring method (component scores, difference scores)}; and estimation method [13–15].
\end{displayquote}

\textbf{4. The authors discuss both classical test theory and generalizability theory - it may be helpful to quickly define each or otherwise provide context for these terms and the assumptions they make.}

Due to word limit constraints, we have elected not to expand on the history and assumptions of classical test theory. Instead, we now cite Allen \& Yen's \cite{allen2001introduction} seminal introduction to classical test theory and measurement theory.

\textbf{5. The authors discuss generative or statistical models as a way to improve reliability at the end of the section on calculating reliability, but I don't see a discussion of this in the section about improving reliability - this topic may be more useful in that section. In particular, models of underlying processes can be more reliable than relying on difference scores, e.g. PMID: 34252724 \cite{weigard2021cognitive}}.

We thank the reviewer for this suggestion. We were deliberate in choosing to place our discussion of hierarchical and generative models in the section discussing calculating reliability. Our motivations for doing so were twofold. First, we did not want to devote much space to hierarchical models in the context of task reliability as this topic has received extensive treatment elsewhere \cite{haines2020learning, Rouder2019-am, chen2021trial}. Second, we instead wanted to primarily focus on principles of experiment design that impact task reliability, which has received less cohesive treatment. We have attempted to make our intentions for the manuscript more clear in the introduction. The text now reads (page 4, line 14-16):

\begin{displayquote}
    The purpose of the current article is to provide a narrative review of approaches to improve task-measure reliability. The article is divided into two main parts. First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks. \textit{This section is not intended to be exhaustive, but rather to point to other publications where these topics have already been discussed at length.} Then, we introduce a taxonomy of approaches to improve the reliability of cognitive-task measures through experiment design and analysis, using concrete examples from the published literature. We hope that this article can serve as a helpful guide for experimenters designing a new task, improving an existing task, or refining their scoring methods to achieve sufficient reliability for use in individual-differences research.
\end{displayquote}

Separately, we thank the reviewer for bringing Weigard et al. to our attention. We now cite their work in the section discussing the use of alternatives to difference scores (page 13, lines 29-31):

\begin{displayquote}
    Another approach is to identify alternative measures of task performance. For example, intra-individual response time variability \textit{and cognitive efficiency have been identified as correlates of executive control that can be measured reliably [80, 81] and are altered in psychopathology [82, 83].}
\end{displayquote}

\textbf{Suggested additions:}

\textbf{1. Page 5, starting line 9 - the overview of pros and cons of different types of internal consistency measurements is helpful. For recommendations, why not randomly assign trials to halves? This does not necessarily require many permutations. Using blocks also has drawbacks: it assumes the task can be split up into several independent blocks of trials, and with few blocks or influential early blocks, has many of the same drawbacks as splitting first vs. second half of trials.}

We thank the reviewer for this comment. We do not want to recommend a single random assignment of trials, as this runs the risk of, by chance, underestimating reliability (if trials in each split are not approximately tau equivalent) or overestimating reliability (if trials in each split exhibit correlated errors) \cite{Parsons2019-jw, Pronk2022-mq}. Instead, we echo others' recommendations of a permutation-based approach that uses many random partitions. Indeed, while we agree that using blocks has drawbacks, they are largely unavoidable for certain types of cognitive tasks (e.g. learning tasks with stationary reward distributions), hence we left these in as well.

\textbf{2. Page 6, starting line 43 - consider adding PMID: 34688897 as citation.}

We thank the reviewer for bringing this article to our attention. We now cite Chen et al. (2021) \cite{chen2021trial} at multiple points throughout the revised manuscript.

\textbf{3. Page 6, starting line 5 - consider adding PMID: 28864865 as citation.}

We thank the reviewer for bringing this article \cite{howlett2017effect} to our attention. However, we are uncertain about the reasons why the reviewer suggested we cite it. The study in question concerns neither individual differences nor task reliability. Though the study does find evidence of practice effects, they are not so severe as to cause range restriction.

\textbf{4. Section 3.1.4 - the authors suggest that online samples produce more variable data, but empirical results generally show that in-person and online studies do not meaningfully differ (e.g., https://doi.org/10.1525/collabra.17213; PMID: 34267650; PMID: 34267650; PMID: 32883156). \cite{nussenbaum2020moving, ito2021validating, chaytor2021construct}}

We thank the reviewer for this thoughtful point, but we do not agree that the published literature generally shows that in-person and online studies do not meaningfully differ. Gillan \& Rutledge \cite{gillan2021smartphones} recently noted that they required a 30\% increase in sample size (from N=461 to N=670) to reliably detect the same association between compulsivity and model-based planning in online vs in-person participants. Moreover, Nussenbaum and colleages \cite{nussenbaum2020moving} actually reported a similar finding; they found that they required a 40\% increase in sample size (from N=45 to N=63) to reliably detect the same association between age and model-based planning in online vs. in-person participants. Although Ito and colleagues \cite{ito2021validating} do not find substantial differences in behavior on average for in-person vs. online settings, they do not compare variability in performance; moreover, only 23 out of their 83 participants completed the study in-person, such that their study is adequately powered to detect only the largest effects. Finally, Chaytor and colleagues \cite{chaytor2021construct} compare two different neuropsychological test batteries between in-person (WAIS) vs. online (TestMyBrain) settings, confounding task and testing context. In short, we believe that the current empirical evidence, if anything, better supports the hypothesis that online data are more variable. As such, we believe we are justified to state, ``experimenters should take special care to ensure that an increase in between-participants variance is not offset by a concomitant increase in measurement noise.''

\textbf{5. An additional consideration of online studies, particularly when using common tasks, is that participants may not be naïve: \textit{Stewart, N., Ungemach, C., Harris, A. J. L., Bartels, D. M., Newell, B. R., Paolacci, G., \& Chandler, J. (2015). The average laboratory samples a population of 7,300 Amazon Mechanical Turk workers. Judgment and Decision Making, 10(5), 479-491} \cite{stewart2015average}.}

We thank the reviewer for this important point. We now address participant nonna\"ivete in the discussion of recruiting participants online. The text now reads (beginning on page 8, line 39):

\begin{displayquote}
    \textit{Separately, online participants may be more familiar with particular experimental paradigms due to previous exposure [49], which may attenuate between-participants performance variability for the reasons previously mentioned (e.g., practice effects). Thus, researchers running experiments online may want to alter task paradigms so that they appear less similar to preexisting 8 versions and/or limit the recruitment of highly-experienced participants [50].}
\end{displayquote}

\textbf{6. The discussion is too brief and needs a summary of findings and recommendations - what can people do to improve reliability when designing a new task, modifying an existing task, or using a task that is already developed? What are remaining issues in reliability of cognitive tasks and future directions?}

We thank the reviewer for this suggestion. Unfortunately, due to word limit constraints, we are unable to provide a substantive discussion section (though we have added an additional paragraph discussing when a task measure is ``reliable enough''; see Reviewer 2, point 1).

\section*{Reviewer \#5}

\textbf{Zorowitz and Niv thoroughly reviewed the definitions of reliability and ways to increase task reliability. Authors suggested two major ways for improving task reliability, which are increasing between-participant variance and decreasing measurement noise. Authors provided theoretical and mathematical formulations to explain why each approach would be beneficial for increasing task reliability. I think this is an excellent comprehensive guideline for interpreting and improving the reliability of task reliability.}

\textbf{Major suggestions:}

\textbf{1. In the first paragraph in section 3.1.3. (p.7), the authors mention that amplifying the magnitude of an experimental effect could improve reliability. Kucina et al. achieved this by increasing the task demand. I felt that this approach might cause the floor effect discussed in this paper, which would ironically decrease the range of responses. This might be related to the reliability paradox, as an attempt to increase the magnitude of an experimental effect decreases the between-participant variability. Are there studies that address this issue? A careful calibration of task difficulty as discussed in the following paragraph would help resolve this problem.}

We thank the reviewer for this point and wish to clarify a misunderstanding. Kucina et al. measure conflict effects via response time, not response accuracy. As such, the issue of floor effects due to increased task demands is obviated. To prevent further reader confusion, we have made clear that those authors measured response times. The text now reads:

\begin{displayquote}
    For example, Kucina and colleagues [26] investigated the reliability of cognitive conflict effects \textit{(as measured by response time)} in new versions of several standard cognitive-control tasks (e.g., Stroop, Flanker, Simon) that amplified cognitive interference via two task design features.
\end{displayquote}

\textbf{2. In the first paragraph in section 3.2.2 (p.9), the authors suggest we use the most discriminating stimuli. It would be worth mentioning the methods for finding such stimuli. Optimal experimental design methods often aim to select stimuli that are most likely to discriminate between participants (or between models). For example, as briefly discussed by authors, adaptive design optimization selects the design that reduces uncertainty about the parameters in a target computational model to the greatest extent. If the model parameters properly reflect a cognitive process of interest, the stimuli that improves the precision of the parameter estimates are likely to be the stimuli that are most discriminating.}

We thank the reviewer for this suggestion. In Section 3.2.2, we now cite Embretson \& Reise \cite{embretson2013item} who provide the gold-standard introduction to item response theory (IRT) for psychologists. IRT methods are ideally suited for identifying the most discriminating stimuli. It is worth noting that adaptive design optimization (ADO) methods do not necessarily identify the most discriminating stimuli (defined here, in line with item response theory, as stimuli for which participants with high ability consistently respond in one way [e.g., make the correct response] while participants with low ability consistently respond in another way [e.g., make an incorrect response]). If an ADO method selects a stimulus based only on the expected response, but not the variability around that expectation, then an ADO method may not necessarily select the most discriminating stimulus.

\printbibliography

\end{document}

### 2023-01-09 manuscript/main.tex
manuscript revisions + letter to reviewers

@@
 \usepackage[a4paper, margin=0.8in]{geometry}
+\usepackage[backend=biber,style=nature]{biblatex}
 \usepackage{graphicx}
@@
 \usepackage{authblk}
-\usepackage{hyperref}
-\usepackage[backend=biber,style=nature]{biblatex}
+\usepackage{hyperref}           % format hyperlinks
+\usepackage[pagewise]{lineno}   % add line numbers
+\usepackage{xcolor,soul}        % highlight formatting
 
@@
 \renewcommand*{\bibfont}{\small}
+
+% define highlight style
+\colorlet{usercolorname}{yellow!00}    % use for public version
+% \colorlet{usercolorname}{yellow!30}    % use for reviewer version
+% \linenumbers                           % use for reviewer version
+\sethlcolor{usercolorname}
 
@@
 \author[1,*]{Samuel Zorowitz}
-\author[1,2]{Yael Niv}
+\author[1,2,*]{Yael Niv}
 \affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
 \affil[2]{Department of Psychology, Princeton University, USA}
-\affil[*]{Corresponding author {\small (zorowitz@princeton.edu)}}
+\affil[*]{Corresponding authors (zorowitz@princeton.edu, yael@princeton.edu)}
 
@@
 
-\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena across individuals.
-However, many recent studies have found that task measures exhibit poor reliability, which hampers their utility for individual-differences research.
+\abstract{Cognitive tasks are capable of providing researchers with crucial insights into the relationship between cognitive processing and psychiatric phenomena.
+However, many recent studies have found that task measures exhibit poor reliability, which hampers their usefulness for individual-differences research.
 Here we provide a narrative review of approaches to improve the reliability of cognitive task measures.
 First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks.
-Then, we introduce a taxonomy of approaches for improving task reliability.
+Then, we introduce a taxonomy of experiment design and analysis strategies for improving task reliability.
 Where appropriate, we highlight studies that are exemplary for improving the reliability of specific task measures.
@@
 
-% P1: Using valid and reliable measures of behavior is crucial for advancing biological psychiatry.
+\hl{Cognitive tasks hold great promise for biological psychiatry.} When properly designed, such tasks are capable of isolating and measuring specific cognitive processes.
+Individual differences in performance on cognitive tasks can therefore provide researchers with crucial insights into the cognitive processes underlying psychiatric phenomena.
+Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{spiegel2021relations} and cognitive decline \cite{hartshorne2015does}.
+\hl{Cognitive tasks, then, have the potential to be invaluable tools for refining our understanding of psychiatric symptoms and syndromes.} For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.
 
-Cognitive tasks have already demonstrated, and continue to hold, tremendous promise for biological psychiatry.
-Individual differences in performance on cognitive tasks provide researchers with crucial insights into the relationship between cognitive processing and behavioral phenomena in the clinic and in everyday life (e.g.
-reward sensitivity in a monetary-incentivized task and anhedonia in depression).
-Elsewhere in psychology, cognitive tasks have been useful in predicting important outcomes such as academic achievement \cite{spiegel2021relations} and cognitive decline \cite{hartshorne2015does}.
-Cognitive tasks, then, are invaluable tools for building and refining our understanding of psychiatric symptoms and syndromes.
-For a cognitive task to be useful in this regard, however, it must possess sufficient measurement properties.
+We define a cognitive task as any experimental paradigm that measures behavior in order to make inferences about one or more cognitive processes (e.g., Stroop task, delay discounting task, reversal-learning task).
+Cognitive-task measures of behavioral performance can be descriptive (e.g., proportion correct responses, average response time) or model-based (e.g., drift rate in evidence accumulation models).
+The psychometric quality of a measure can be summarized by three key properties: \emph{discriminatory power}, \emph{validity} and \emph{reliability} \cite{kline2015handbook}.
+The discriminatory power of a task measure describes its ability to measure variability in participants' performance.
+This is a necessary property of tasks used to study individual differences; where there is no variation in performance, there are no individual differences to study.
+The validity of a task measure concerns whether it actually measures what it intends to measure.
+Finally, the reliability of a task measure characterizes the degree to which it consistently measures some feature of participants.
+\hl{That is, a task measure is reliable if, assuming participants have not changed, it produces the same scores, or the same ordering of scores, for participants within a single testing session or across multiple testing sessions.} This review focuses on task-measure reliability.
 
-% P2: Definitions + Three key properties of behavioral measures: discriminability, reliability, and validity
-
-We define a cognitive task as any experimental paradigm that measures behavior in order to make inferences about one or more cognitive processes (e.g.
-Stroop task, delay discounting task, reversal learning task).
-Cognitive-task measures for characterizing individual differences are descriptive or model-based indicators of behavioral performance such as proportion correct, average response time, and learning rate.
-The psychometric quality of such a measure can be summarized by three key properties: \emph{discriminating power}, \emph{validity} and \emph{reliability} \cite{kline2015handbook}.
-The discriminatory power of a task measure describes its ability to measure variability in participants' performance.
-This is a necessary property of tasks to be used for studying individual differences; where there is no variation in performance, there are no individual differences to study.
-The validity of a task measure concerns whether it actually measures what it intends to measure.
-Finally, the reliability of a task measure characterizes the degree to which it consistently rank-orders participants across measurement instances.
-Put another way, a task measure is reliable if, assuming participants have not changed, it produces the same score for each participant over time.
-The focus of this review is task reliability.
-% P3: definition and Importance of reliability
 \subsubsection*{The formal definition of reliability}
 
-In classical test theory, the variance in observed scores on a task measure $x$ is the sum of the true score variance $\sigma^2_T$, reflecting real individual differences in the latent construct of interest, and measurement error $\sigma^2_E$, i.e., $\sigma^2_x = \sigma^2_T + \sigma^2_E$.
+In classical test theory \cite{allen2001introduction}, the variance in observed scores on a task measure $x$ is the sum of the true score variance $\sigma^2_T$, reflecting real individual differences in the latent construct of interest, and measurement error $\sigma^2_E$, i.e., $\sigma^2_x = \sigma^2_T + \sigma^2_E$.
 The reliability of a measure is defined as the proportion of variance attributable to the true score variance relative to total variance: $\rho_{xx'} = \frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$.
-Thus, reliability quantifies the magnitude of individual differences relative to the noisiness of a task measure; the larger the reliability of a task measure, the more it reflects true individual differences rather than reflecting noise.
+Thus, reliability quantifies the magnitude of individual differences relative to the noisiness of a task measure; the larger the reliability of a task measure, the more it reflects true individual differences rather than noise.
 Reliability is therefore a prerequisite for validity: an unreliable task measure reflects measurement error and not the construct of interest.
-If this were not reason enough to care about reliability, the observed correlation between two measures (e.g.
-task performance and self-reported symptom score) is bounded by their individual reliabilities \cite{Spearman1904-mo}:
+If this were not reason enough to care about reliability, the observed correlation between two measures (e.g., task performance and self-reported symptom score) is bounded by their individual reliabilities \cite{Spearman1904-mo}:
 \begin{equation}
@@
 
-% P4: Reliability is inherently context-dependent: it is a function of the population, task, and testing context
-
-To further complicate matters, the reliability of a task measure is not absolute -- it reflects interactions between task design, the participants, and the context in which the task is administered.
-Indeed, task reliability can vary as a function of parameters of the experiment such as the specific stimulus set, number of trials, time limits \cite{paap2016role, cooper2017role}; sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in clinic, online); response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); and scoring or estimation method \cite{Rouder2019-am, haines2020learning}.
+To further complicate matters, the reliability of a task measure is not absolute -- it reflects interactions between the task design, the participants, and the context in which the task is administered.
+Indeed, task reliability can vary as a function of experiment parameters (specific stimulus set, number of trials, time limits \cite{paap2016role, cooper2017role}); sample populations (healthy adults, children, psychiatric patients \cite{arnon2020current, cooper2017role}); testing locations (in clinic, online); response modality (desktop, smartphone, virtual reality \cite{pronk2022can, bruder2021reliability}); \hl{scoring method (component scores, difference scores);} and estimation method \cite{Rouder2019-am, haines2020learning, chen2021trial}.
 For example, a cognitive task originally designed for use with an adult population may prove too difficult for children.
@@
 Experimenters therefore cannot assume the reliability of a task measure is constant.
-Ideally researchers would investigate and report the reliability of task measures as part of any individual differences research or, at the very least, after having made changes to a task or scoring procedure or when administering the task to new sample populations or in new testing contexts.
-
-% P5: Task reliability lower on average
+At the very least, researchers should evaluate reliability after having made changes to a task or scoring procedure, or when administering the task to new sample populations or in new testing contexts.
+Ideally, researchers would investigate and report the reliability of task measures as part of any individual differences research.
 
@@
 This is the exact opposite of what is desirable for individual differences research, where between-participants variance is essential to achieving reliable task measures.
-Thus, the tendency in biological psychiatry to adopt the most prominent tasks in experimental psychology---the ones that most reliably demonstrate a behavioral effect---may actually hamstring our efforts to study individual differences.
+\hl{For example, the Stroop effect is one of the most robust effects in experimental psychology; virtually everyone shows a Stroop effect} \cite{haaf2017developing}.
+\hl{However, in part due to this fact, between-participants variance on the Stroop effect is often limited} \cite{rouder2019most}.
+Thus, the tendency in biological psychiatry to adopt the most prominent tasks in experimental psychology---the ones that most reliably demonstrate a behavioral effect---may actually hamstring efforts to study individual differences.
 
-% P6: Purpose of paper: narrative review of task reliability
-
-Regardless, we do not believe that task measures are inherently less reliable than self-report measures, and that pessimism about task-based individual-differences research is warranted.
-It is possible to (re)design tasks to achieve good reliability, even to the high levels dictated by conventional standards \cite{waltmann2022sufficient, sullivan2022enhancing, kucina2022solution, snijder2022psychometric}.
+Regardless, we do not believe that task measures are inherently less reliable than self-report measures, or that pessimism about task-based individual-differences research is warranted.
+It is possible to (re)design tasks to achieve good reliability, even to the high levels dictated by conventional standards \cite{mclean2018towards, kucina2022solution, snijder2022psychometric, rey2019executive}.
 The purpose of the current article is to provide a narrative review of approaches to improve task-measure reliability.
@@
 First, we review methods of calculating reliability and discuss some nuances that are specific to cognitive tasks.
-Then, we introduce a taxonomy of approaches for improving task reliability using concrete examples from the published literature.
-We hope that this article can serve as a helpful guide for experimenters who wish to design a new task, or improve an existing task, to achieve sufficient reliability for use in individual-differences research.
+\hl{This section is not intended to be exhaustive, but rather to point to other publications where these topics have already been discussed at length.} Then, we introduce a taxonomy of approaches to improve the reliability of cognitive-task measures through experiment design and analysis, using concrete examples from the published literature.
+We hope that this article can serve as a helpful guide for experimenters designing a new task, improving an existing task, \hl{or refining their scoring methods} to achieve sufficient reliability for use in individual-differences research.
 
@@
 Even for cross-sectional experiments, where stability over time is of no scientific interest, test-retest reliability is a useful index.
-This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant, state-dependent factors (e.g.
-current mood, fatigue).
+This is because estimates of internal consistency tend to be inflated due to between-participants variance from construct-irrelevant, state-dependent factors (e.g., current mood, fatigue).
 Given sufficient time between testing sessions, test-retest reliability should be less biased by state factors, and thus a better estimate of the true cross-sectional task-measure reliability.
@@
 Calculating the internal consistency of a task measure is more complicated.
-The most common measure of internal consistency, Cronbach's $\alpha$, requires certain assumptions that are unrealistic for many task data (e.g., equivalence of trials, uncorrelated measurement error).
-As such, internal consistency for task measures is instead usually calculated via split-$K$ reliability, where $K$ indicates the number of splits.
-Split-half reliability ($K=2$) is most common.
-A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be biased when partitions are not equally good measures of a behavior \cite{Green2016-xw, pronk2022can}.
-For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to differences between the partitions resulting from practice, fatigue, or other linear time effects.
-In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may be confounded when behavior across trials is non-independent (i.e., measurement error is correlated across trials).
-Under such conditions, odd-even splitting will artificially inflate the similarity of data across partitions, thereby decreasing estimates of measurement noise and overestimating reliability.
-Therefore, where possible, a permutation-based approach to calculating split-half reliability is recommended \cite{pronk2022can, Parsons2019-jw}.
+\hl{The most common measure of internal consistency is Cronbach's $\alpha$, which is a function of the average correlation across all unique pairs of trials.
+However, Cronbach's $\alpha$ is an accurate measure of reliability only under assumptions that are unrealistic for many tasks (e.g., equivalence of trials, uncorrelated measurement error;} \cite{Green2016-xw, Parsons2019-jw}).
+\hl{As such, internal consistency for task measures is instead usually calculated via split-half reliability, where reliability is estimated after trial data have been divided into two halves.} A critical challenge in calculating split-half reliability is in deciding how to partition the data, as estimates of reliability may be also biased if the data partitions violate either of the two above assumptions (for detailed discussion, see \cite{Green2016-xw, Pronk2022-mq}).
+For example, first-second splitting (i.e., partitioning the data into the first and second halves of an experiment) may underestimate reliability due to nonequivalence of the two partitions resulting from practice, fatigue, or other linear time effects.
+In contrast, odd-even splitting (i.e., partitioning the data into odd and even trials) may cause bias when behavior across trials is non-independent (i.e., measurement error is correlated across trials),  artificially inflating the similarity of data across partitions and thereby decreasing estimates of measurement noise and overestimating reliability.
+Therefore, where possible, a permutation-based approach to calculating split-half reliability is recommended \cite{Pronk2022-mq, Parsons2019-jw}.
 Here, reliability is averaged across many thousands of random partitions of the data into halves.
-For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the model for each new subset of the data.
+\hl{(Insofar that Cronbach's $\alpha$ is analytically equivalent to the average of all possible split-half reliability estimates} \cite{cronbach1951coefficient}, \hl{permutation-based split-half reliability provides an approximation to Cronbach's $\alpha$ while avoiding its problematic assumptions.)} For task measures derived from cognitive models, however, it may be prohibitively computationally intensive to employ such an approach due to the need to re-estimate the model for each new subset of the data.
 Moreover, for learning tasks that are commonly used in computational psychiatry, cognitive-model based task measures cannot be estimated from only a subset of the trials due to inherent non-independence of task behavior across trials.
@@
 
-As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning}.
-This is because such summary scores are contaminated by trial-level noise that, in the absence of sufficiently large number of trials, increases the measurement error (and thus diminishes reliability).
-Instead, it is often preferable to use generative or statistical models that account for trial-level variability in behavior.
-By directly modeling trial-by-trial variability, estimates of participants' performance are effectively de-noised.
-It is also advisable to use hierarchical models that partially pool data across participants, which in turn further increases the precision of estimates of participants' performance.
-Together, these methods can substantially boost estimates of reliability \cite{Rouder2019-am, haines2020learning}, as multiple studies have demonstrated  \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}.
-For a detailed discussion of generative models in the context of task reliability, see Haines (this issue).
+As a final point, traditional sum or mean score estimates of performance (e.g., proportion correct responses, mean response time) may substantially underestimate task reliability \cite{Rouder2019-am, haines2020learning, chen2021trial}.
+\hl{This is because such summary scores are contaminated by trial-level noise that, in the absence of a sufficiently (possibly prohibitively) large number of trials, increases measurement error (and thus diminishes reliability).
+Instead, it may be preferable to use trial-level hierarchical models in which observations are organized hierarchically (e.g., individuals within a group, trials within an individual) with variability modeled at both levels.
+Hierarchical models exert a pooling or regularization effect on person-level variables, in effect correcting for measurement error and improving estimates of reliability} \cite{Rouder2019-am, haines2020learning, chen2021trial}.
+The benefits of hierarchical models for estimating reliability has been multiply demonstrated \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}, \hl{though see} \cite{rouder2019most} \hl{for discussion of when these benefits may be limited.
+Using statistical models that more accurately characterize the latent data-generating process (e.g., using the shifted log-normal distribution to model response times) may also improve reliability estimates} \cite{haines2020learning, price2019computational}.
+For a detailed discussion of hierarchical and generative models in the context of task reliability, see Haines and colleagues (this issue; \cite{haines2020learning}).
 
@@
 In response, the authors designed a new statistical learning task involving stimulus sequences that ranged more widely in their difficulty to learn.
-Only a minority of participants showed chance-level performance on this new task and, as such, the reliability of proportion correct scores improved.
+Only a minority of participants showed chance-level performance on this new task and, as such, the reliability of proportion correct scores improved \hl{(from $\rho = 0.75$ to $0.88$)}.
 Similarly, in developing an abbreviated working-memory task, Oswald and colleagues \cite{oswald2015development} found that they could remove the easiest trials---those with ceiling level performance---with virtually no change to task reliability.
@@
 For example, Arnon and colleagues \cite{arnon2020current} found that statistical learning tasks developed for adults were too difficult for young children and therefore yielded unreliable discrimination scores in that population.
-Similarly, Kyllonen and colleagues \cite{kyllonen2019general} were motivated to develop a battery of fluid-reasoning measures for highly educated adults due to observed ceiling effects in performance when using preexisting fluid-reasoning tasks in this population.
+Similarly, Kyllonen and colleagues \cite{kyllonen2019general} developed a battery of fluid-reasoning measures for highly educated adults after observing ceiling effects in performance when using preexisting fluid-reasoning tasks in this population.
 
@@
 Practice effects are relatively common for cognitive tasks \cite{hausknecht2007retesting, scharfen2018retest}.
-They might occur due to the attenuation of task-irrelevant nuisance factors (e.g.
-performance anxiety) and/or the learning of task-specific knowledge or strategies.
-Practice effects are not inherently an issue for reliability --- especially if an experimenter is only interested in the consistency, but not the absolute agreement, of participants' performance over time --- but they can become a pernicious issue if they are severe enough to induce ceiling effects.
+They might occur due to the attenuation of task-irrelevant nuisance factors (e.g., performance anxiety) and/or the learning of task-specific knowledge or strategies.
+Practice effects are not inherently an issue for reliability --- especially if an experimenter is only interested in the consistency, but not the absolute agreement, of participants' performance over time --- but they can become a pernicious issue if \hl{they are exhibited differentially across participants} or if they are severe enough to induce ceiling effects.
 For example, Paredes and colleagues \cite{paredes2021psychometric} observed large practice effects on the Pavlovian go/no-go task for short retest intervals (3 days, 14 days), which resulted in poor estimates of test-retest reliability.
@@
 In response, the authors developed a new version of the task that included distractor sequences of bead draws.
-This new design was effective in preventing participants from becoming aware of the target sequence, which in turn resulted in more consistent responding, which improved the reliability of participants' information seeking scores.
+This new design was effective in preventing participants from becoming aware of the target sequence, which in turn resulted in more consistent responding, which improved the reliability of participants' information seeking scores \hl{(from $\rho = 0.62$ to $0.84$)}.
 
-\subsubsection{Enhancing experimental effects}
+\subsubsection{Enhancing experimental manipulations}
 
 The preceding sections described potential threats to between-participants variability, but not approaches to improve it.
-A primary strategy for increasing between-participants variability is to enhance the experimental effect.
-Amplifying the magnitude of an experimental effect (e.g., making a task more challenging, increasing the potency of affect induction) typically increases the range of participants' responses to it.
-For example, Kucina and colleagues \cite{kucina2022solution} investigated the reliability of cognitive conflict effects in new versions of several standard cognitive-control tasks (e.g., Stroop, Flanker, Simon) that amplified cognitive interference via two task design features.
+A primary strategy for increasing between-participants variability is to enhance the experimental \hl{manipulation}.
+Amplifying the strength of an experimental \hl{manipulation} (e.g., making a task more challenging, increasing the potency of affect induction) typically increases the range of participants' responses to it.
+For example, Kucina and colleagues \cite{kucina2022solution} investigated the reliability of cognitive conflict effects \hl{(as measured by response time)} in new versions of several standard cognitive-control tasks (e.g., Stroop, Flanker, Simon) that amplified cognitive interference via two task design features.
 First, they combined multiple sources of cognitive interference in the same task, for example by combining the Stroop and Simon effects to create a ``Stroopon'' task.
 Second, for a subset of trials, they required participants to make multiple responses based on both relevant and irrelevant stimuli attributes.
-Compared to the standard version of these tasks, these experimental manipulations had the effect of increasing task demands, which resulted in greater between-participants variance, and higher reliability as a consequence.
-(See also Snijder and colleagues \cite{snijder2022psychometric} for a similar redesign of classic cognitive-control tasks that improved reliability by increasing proactive control demands.)
+Compared to previous versions of these tasks, these manipulations had the effect of increasing task demands, which resulted in greater between-participants variance and, consequently, \hl{required hundreds fewer trials to achieve a reliability of $\rho = 0.8$.} (See also Snijder and colleagues \cite{snijder2022psychometric} for a similar redesign of classic cognitive-control tasks that improved reliability in part by increasing proactive control demands.)
 
@@
 Of course, this design principle is only helpful to the degree that a researcher knows the average ability level of their participants.
-If this is unknown or poorly characterized, then it is instead desirable to design a task with trials spanning a uniform range of difficulty.
+If this is unknown or poorly characterized, then it is instead desirable to design a task with trials spanning a range of difficulty levels.
 
@@
 Thus, when recruiting online samples, experimenters should take special care to ensure that an increase in between-participants variance is not offset by a concomitant increase in measurement noise.
+\hl{Separately, online participants may be more familiar with particular experimental paradigms due to previous exposure} \cite{chandler2014nonnaivete}, \hl{which may attenuate between-participants performance variability for the reasons previously mentioned (e.g., practice effects).
+Thus, researchers running experiments online may want to alter task paradigms so that they appear less similar to preexisting versions and/or limit the recruitment of highly-experienced participants} \cite{robinson2019tapped}.
 
@@
 Perhaps the most straightforward approach to decreasing measurement error, and thereby increasing reliability, is to increase the number of task trials.
-In generalizability theory, the relationship between reliability and the number of trials defined as:
+The relationship between reliability and the number of trials defined as:
 \begin{equation}
@@
 For example, in the context of an emotion recognition task, stimuli of good discriminability would be those where participants with good emotion-recognition ability consistently correctly identify the displayed emotion while participants with poor ability consistently incorrectly identify the displayed emotion \cite{keutmann2015generating}.
-In contrast, stimuli with poor discriminability --- those for which performance between high- and low-ability participants is indistinguishable --- will lead to more measurement noise and decreased reliability.
-In experiments where stimuli are intended to be unique and distinguishable, improving both the linguistic and visual distinctness of stimuli may prevent confusion and therefore aid reliability \cite{yoo2022importance}.
+In contrast, stimuli with poor discriminability --- those for which performance between high- and low-ability participants is indistinguishable --- will lead to more measurement noise and decreased reliability \cite{embretson2013item}.
+In experiments where stimuli are intended to be unique and distinguishable, improving both the linguistic and visual distinctness of stimuli may prevent participant confusion and therefore aid reliability \cite{yoo2022importance}.
 
 Other design features of an experiment that are specific to the task mechanics may affect reliability.
-Consider for example the dot-probe task, in which participants must disengage attention from a distracting image on one part of the screen in order to identify and respond to the orientation of a pair of dots elsewhere on the screen.
-Dot-bottom trials, where a participant must disengage from a distracting stimulus located at the top of the screen and saccade to the bottom of the screen, are more reliable than dot-top trials \cite{price2015empirical, aday2019extended}.
-This has been explained by suggesting that participants are biased to gaze towards the top of the screen, thus saccading away from the top of the screen requires a stronger level of disengagement, allowing such trials to better measure attentional bias.
+Consider, for example, the dot-probe task, in which participants must disengage attention from a distracting image on one part of the screen in order to identify and respond to the orientation of a pair of dots elsewhere on the screen.
+Dot-bottom trials, where a participant must disengage from a distracting stimulus located at the top of the screen and saccade to the bottom of the screen, are more reliable than dot-top trials \hl{(e.g., dot-bottom: $\rho = 0.33$; dot-top: $\rho = 0.07$)} \cite{price2015empirical, aday2019extended}.
+This has been explained by suggesting that because participants' gazes are biased towards the top of the screen, saccading away from the top requires a stronger level of disengagement.
+Dot-bottom trials may therefore be better measures of attentional bias.
 
@@
 Thus, practice trials can help participants reach a ``steady state'' of responding, thereby reducing the noisiness of their responses across task trials and increasing reliability \cite{alexander2003effects}.
-Practice can take the form of a standalone practice block or by designating as such and discarding (or modeling separately) the first few trials of an experiment \cite{mclean2018towards}, though the latter may not fully allow participants to explore response options during the practice.
+Practice can take the form of a standalone practice block or by designating as such and discarding (or modeling separately) the first few trials of an experiment (e.g, \cite{mclean2018towards}), though the latter may not fully allow participants to explore response options during the practice.
 
@@
 Incorporating (video) game design elements into cognitive tasks can increase participant engagement and motivation \cite{sailer2017gamification}, countering the would-be effects of boredom and fatigue on task reliability.
-For example, Kucina and colleagues \cite{kucina2022solution} cite task gamification as an important factor that contributed to the improved the reliability of their cognitive-control tasks.
-Similarly, Verdejo and colleagues \cite{verdejo2021unified} partially attribute the improved reliability of their impulsivity task battery to gamified task design.
+For example, Kucina and colleagues \cite{kucina2022solution} cite task gamification as an important factor that contributed to the reliability of their cognitive-control tasks.
+Similarly, Verdejo and colleagues \cite{verdejo2021unified} partially attribute the adequate reliability of their impulsivity task battery to gamified task design \hl{(range: $\rho = 0.52 - 0.71$)}.
 
@@
 When parameters from cognitive models are used as indices of participants' task performance, another means to improve measure reliability is to decrease estimation noise.
-The estimation noise of a parameter given an experiment and model can be quantified through simulation \cite{wilson2019ten, palminteri2017importance}.
+The estimation noise of a parameter given an experiment and model can be quantified through simulation studies \cite{wilson2019ten, palminteri2017importance}.
 Here, an experimenter generates artificial data for the experiment using representative model parameters and then attempts to recover the model parameters by fitting the model to the simulated data.
@@
 A related approach is to use adaptive experimental designs \cite{myung2013tutorial}, where the trials of an experiment are designed in real time so as to present each participant with stimuli or trial types that are matched to their particular response patterns or ability levels.
-Though undoubtedly a more complex experimental design, adaptive experiments have the advantage of selecting the most informative trials for resolving the ability or preference level of a participant (e.g.
-as measured by a cognitive model).
+Though undoubtedly a more complex experimental design, adaptive experiments have the advantage of selecting the most informative trials for resolving the ability or preference level of a participant (e.g., as measured by a cognitive model).
 Adaptive designs have been successfully used in cognitive research, for example to study working memory \cite{gonthier2018measuring} and delay discounting \cite{ahn2020rapid}.
@@
 Parameter estimation may be further improved by leveraging additional information.
-For example, latent variables may be more accurately measured through the inclusion of demographic variables or other covariates \cite{Curran2016-bt} that, if associated to model parameters, can aid in resolving parameter estimates.
+For example, latent variables may be more accurately measured through the inclusion of demographic variables or other covariates that, if associated to model parameters, can aid in resolving parameter estimates \cite{Curran2016-bt}.
 An extension of this idea is to utilize joint modeling of dependent variables; that is, to design models where multiple observed trial-level variables are predicted simultaneously.
@@
 Difference scores subtract a measure of a participant's performance in one condition from their performance in another.
-An example is the classic Stroop interference effect: the average reaction-time difference between congruent and incongruent trials.
-Difference scores are commonly used because they allow experimenters to isolate particular cognitive processes (e.g.
-processing cognitive conflict) while controlling for other sources of variance (e.g.
-perceptual processing, motor ability) through the subtraction of conditions that share that variance.
+An example is the Stroop interference effect, calculated as the average reaction-time difference between congruent and incongruent trials.
+Difference scores are commonly used because they allow experimenters to isolate particular cognitive processes (e.g., processing cognitive conflict) while controlling for other sources of variance (e.g., perceptual processing, motor ability) through the subtraction of conditions that share that variance.
 The challenge is that the reliability of a difference score is a function of the reliability of each of its components \textit{and} the correlation between the components:
@@
 (It is also worth noting that component variances essentially function as weights in determining the overall reliability, such that the difference-score reliability reflects more the reliability of the component measure with the larger variance; Figure \ref{fig:fig03}B.) This speaks to the advantage of increasing the between-participants variability of a measure of performance in one experiment condition without increasing performance variability in a second condition.
-This may explain why Kuchina and colleagues improved the reliability of their difference score by making only the incongruent trials more difficult \cite{kucina2022solution}.
+This may explain how Kuchina and colleagues improved the reliability of their difference score measures after making only the incongruent trials more difficult \cite{kucina2022solution}.
 
@@
 Rey-Mermet and colleagues \cite{rey2019executive} provide an interesting example in the context of executive control.
-In typical executive control tasks, response times on congruent and incongruent trials are highly correlated, reflecting shared variance from conflict-irrelevant processes including baseline processing speed (e.g.
-perceptual processing, motor speed) and performance strategies (e.g.
-individual differences in speed-accuracy preferences; \cite{draheim2019reaction}).
+In typical executive control tasks, response times on congruent and incongruent trials are highly correlated, reflecting shared variance from conflict-irrelevant processes including baseline processing speed (e.g., perceptual processing, motor speed) and performance strategies (e.g., individual differences in speed-accuracy preferences; \cite{draheim2019reaction}).
 Rey-Mermet and colleagues designed a number of ``response-deadline'' executive control tasks where participants had a limited amount of time to respond during a trial.
@@
 It also controls for individual differences in strategy, as regardless of whether a participant was biased towards speed or accuracy, inefficient executive control would result in lower accuracy.
-With this calibration procedure, the researchers found that the reliability of an accuracy difference score (incongruent - congruent) was as good or better than what had previously been reported for executive control tasks.
+With this calibration procedure, the researchers found that the reliability of an accuracy difference score (incongruent minus congruent, \hl{which ranged from $\rho = 0.58$ to $\rho = 0.91$}) was as good or better than what had previously been reported for executive control tasks.
 Thus, controlling for shared variance across measures---that is, purifying measures---can help to improve task reliability.
@@
 Draheim and colleagues \cite{draheim2019reaction} provide an in-depth review of alternatives to difference scores in the context of response-time measures, though much of their discussion is applicable to difference scores in general.
-One possibility is to simply use the component measures (e.g.
-performance on incongruent trials in a Stroop task alone).
+One possibility is to simply use the component measures (e.g., performance on incongruent trials in a Stroop task alone).
 Of course, because component scores will be contaminated other sources of variance, such as baseline performance, interpreting component scores should be done with caution.
 Another approach is to identify alternative measures of task performance.
-For example, intra-individual response time variability has been identified as a correlate of executive control that can be measured reliably \cite{saville2011stability} and is altered in psychopathology \cite{kofler2013reaction}.
+For example, intra-individual response time variability and \hl{cognitive efficiency have been identified as correlates of executive control that can be measured} reliably \cite{saville2011stability, weigard2021cognitive} and are altered in psychopathology \cite{kofler2013reaction, heathcote2015decision}.
 
@@
 We have briefly reviewed issues and research regarding the reliability of cognitive-task measures.
-Specifically, we introduced a taxonomy of approaches for improving cognitive task-measure reliability, highlighting example studies that improved the reliability of specific task measures.
+Specifically, we introduced a taxonomy of experiment design and analysis strategies for improving the reliability of cognitive-task measures, highlighting exemplary studies that have successfully implemented such approaches.
 We hope we have made clear the importance of calculating (and reporting) the reliability of task measures intended for use in psychiatric research.
@@
 
-We emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures.
+\hl{We conclude with two important points.
+First, although we have discussed the importance of task reliability, we have largely avoided the question of when a task measure is ``reliable enough''.
+Though it is tempting to fall back on conventional cutoffs (e.g., $\rho \geq 0.7$), what constitutes sufficient reliability in actuality will depend on the goal(s) of the researcher.
+If the goal is to detect a significant individual-differences correlation, such as between a task measure and self-reported symptom measure, then a task measure with ``unacceptable'' reliability by conventional standards may suffice (e.g., if a researcher has the resources to collect a sample large enough to be adequately powered to detect a correlation at the attenuated magnitude).
+On the other hand, if a researcher intends to estimate an individual-differences correlation with high precision, or use a task measure in a high stakes setting (e.g., treatment selection for an individual patient), then high reliability may be required.
+We cannot overstate the value of simulation studies} (e.g., \cite{rouder2019most}) \hl{for researchers trying to determine what level of reliability is required to meet their goals and risk preferences.}
+
+Second, we would like to emphasize that reliability is but one of many important considerations in the design and evaluation of cognitive task measures.
 Task measures may be reliable but show poor convergent validity \cite{snijder2022psychometric, eckstein2022interpretation}, raising questions about whether they actually measure the constructs they are intended to measure.
 Similarly, task measures may be reliable but exhibit poor ecological validity \cite{Steiner2021-oq}, thus being poor proxies for cognition in real-world settings.
-Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables of interest (e.g.
-symptoms, treatment response).
-Finally, there are many other practical considerations (e.g.
-task duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully in the clinic or in naturalistic environments at scale \cite{germine2021toward}.
+Task measures may also be reliable but show poor predictive validity \cite{verdejo2021unified}, explaining little (unique) variance in other variables of interest (e.g., symptoms, treatment response).
+Finally, there are many other practical considerations (e.g., task duration, engagement, accessibility) to ensure cognitive tasks are able to be deployed successfully in the clinic or in naturalistic environments at scale \cite{germine2021toward}.
 
-Despite the challenges of making task-measures reliable and valid, we are optimistic about their current and future use in biological psychiatry.
-We firmly believe that, with further research efforts towards developing, documenting, and sharing reliable task paradigms, our field can make great strides towards understanding, predicting, and ultimately relieving psychiatric illness.
+Despite the challenges of making task measures reliable and valid, we are optimistic about their current and future use in biological psychiatry.
+We believe that, with further efforts towards developing, documenting, and sharing reliable task paradigms, our field can make increased strides towards understanding, predicting, and ultimately relieving psychiatric illness.
 
@@
 
-We thank Rachel Bedder for helpful feedback on this manuscript.
+We thank Rachel Bedder and Felix Jan Nitsch for helpful feedback on this manuscript.
 This project was made possible with support from the National Center for Advancing Translational Sciences (NCATS), a component of the National Institute of Health (NIH), under award number UL1TR003017a and a National Science Foundation Graduate Research Fellowship.
@@
 First, we obtained the predicted gender of the first and last author of each reference by using databases that store the probability of a first name being carried by a woman \cite{dworkin2020extent}.
-By this measure (and excluding self-citations to the first and last authors of our current paper), our references contain 8.92\% woman(first)/woman(last), 19.37\% man/woman, 21.92\% woman/man, and 49.79\% man/man.
+By this measure and excluding self-citations to the first and last authors of our current paper), our references contain 8.3\% woman(first)/woman(last), 14.7\% man/woman, 18.6\% woman/man, and 58.4\% man/man.
 This method is limited in that a) names, pronouns, and social media profiles used to construct the databases may not, in every case, be indicative of gender identity and b) it cannot account for intersex, non-binary, or transgender people.
 Second, we obtained predicted racial/ethnic category of the first and last author of each reference by databases that store the probability of a first and last name being carried by an author of color \cite{ambekar2009name, sood2018predicting}.
-By this measure (and excluding self-citations), our references contain 4.81\% author of color (first)/author of color(last), 15.59\% white author/author of color, 16.92\% author of color/white author, and 62.68\% white author/white author.
+By this measure (and excluding self-citations), our references contain 4.7\% author of color (first)/author of color(last), 15.5\% white author/author of color, 18.0\% author of color/white author, and 61.8\% white author/white author.
 This method is limited in that a) names and Florida Voter Data to make the predictions may not be indicative of racial/ethnic identity, and b) it cannot account for Indigenous and mixed-race authors, or those who may face differential biases due to the ambiguous racialization or ethnicization of their names.

### 2023-01-13 manuscript/main.tex
fix typos; update to 'in press' status

@@
 \author[1,*]{Samuel Zorowitz}
-\author[1,2,*]{Yael Niv}
+\author[1,2]{Yael Niv}
 \affil[1]{Princeton Neuroscience Institute, Princeton University, USA}
 \affil[2]{Department of Psychology, Princeton University, USA}
-\affil[*]{Corresponding authors (zorowitz@princeton.edu, yael@princeton.edu)}
+\affil[*]{Corresponding author (zorowitz@princeton.edu)}
 
@@
 Hierarchical models exert a pooling or regularization effect on person-level variables, in effect correcting for measurement error and improving estimates of reliability} \cite{Rouder2019-am, haines2020learning, chen2021trial}.
-The benefits of hierarchical models for estimating reliability has been multiply demonstrated \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}, \hl{though see} \cite{rouder2019most} \hl{for discussion of when these benefits may be limited.
+The benefits of hierarchical models for estimating reliability have been multiply demonstrated \cite{snijder2022psychometric, sullivan2022enhancing, brown2020improving, waltmann2022sufficient}, \hl{though see} \cite{rouder2019most} \hl{for discussion of when these benefits may be limited.
 Using statistical models that more accurately characterize the latent data-generating process (e.g., using the shifted log-normal distribution to model response times) may also improve reliability estimates} \cite{haines2020learning, price2019computational}.
